---
title: "Qualitative Summaries"
author: "YOUR NAME HERE"
date: "`r Sys.time()`"
output: html_document
---

Consider the Flight Delays dataset discussed in Chapter 1 of the Chihara & Hesterberg textbook, which contains information on selected flights departing from LaGuardia airport in 2009. Complete all problems below using that dataset. Create all plots using `ggplot2` and be sure to add an informative title and labels for the x and y axes.

```{r}
# Load any necessary packages here. Suppress messages for this code chunk.
```


### Problem 1

Create frequency tables for all of the qualitative variables.

```{r}

```


### Problem 2

Create a barplot to display the frequency of each destination.

```{r}

```


### Problem 3

Create a barplot to show the relative frequency of each destination.

```{r}

```


### Problem 4

Use the information found [HERE](http://www.cookbook-r.com/Manipulating_data/Changing_the_order_of_levels_of_a_factor/) to reorder the levels of the departure time variable so they start at 4am and go in order around the clock. Then create a table that shows cumulative frequency of flights through the day and a barplot that shows the fraction of flights within each time period that are delayed by thirty minutes or more.

```{r}

```


### Problem 5

For two qualitative variables of your choice, create a table and a plot to summarize their relationship. What is one question your summaries might help answer?

```{r}

```


<hr>

END!!!


