---
title: "Dataset Structure Activity"
author: "Instructor: Jill E. Thomley (thomleyje@appstate.edu)"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    highlight: textmate
    theme: cosmo
---

<hr>

Read in each dataset using the `read.csv` function and use the following functions to explore its structure. You can find all the data on our [class dataset list](https://stat-jet-asu.github.io/Datasets/DatasetList.html). The flight delays and Verizon datasets are from the textbook, the rest are instructor datasets.

* `str`
* `glimpse`
* `summary`
* `head`


```{r, message=FALSE, warning=FALSE}
require(dplyr)
```


### Annual Inflation Data

```{r}

```


### Anthropometric Data

```{r}

```


### Flight Delays Data

```{r}

```


### Mauna Loa CO2 Data

```{r}

```


### STT3850 Personality Data

```{r}

```


### Popular Kids Data

```{r}

```


### Powerball Jackpot Data

```{r}

```


### Smallpox Data

```{r}

```


### Verizon Repair Times Data

```{r}

```


### Watauga County Schools Weather Data

```{r}

```



