---
title: "Hypothesis Testing: Comparing Two Groups"
author: "YOUR NAME HERE"
date: "`r Sys.time()`"
output: html_document
---

Consider the Flight Delays dataset discussed in Chapter 1 of the Chihara & Hesterberg textbook, which contains data on selected flights departing from LaGuardia airport in 2009. Complete all problems below using that dataset. 

```{r}
# Load any necessary packages here. Suppress messages for this code chunk.
```


### Problem 1

Conduct a traditional t-test and a permutation resampling test to determine whether the average delay times of the two airlines differ. 

```{r}

```


### Problem 2

Conduct a traditional z-test and a permutation resampling test to determine whether United Airlines has a greater proportion of flights delayed by 30 minutes than American Airlines. 

```{r}

```


### Problem 3

Conduct a traditional F-test and permutation resampling test to determine whether the variance in flight delay times is different between the two carriers. Unlike means and proportions, the test statistic used for a test of two variances is the _ratio_ of the variances, not the difference.

```{r}

```


<hr>

END!!!


