---
title: "Last Day Activity"
author: "YOUR NAME"
date: "6/27/2018"
output: html_document
---

### PROBLEM 1

In class, we worked problems 6-10 in the Sampling Distributions Part 1 practice quiz. Verify your answers using simulation.

```{r}

```

### PROBLEM 2

Consider the Verizon repair times dataset in the book. For the CLEC group...

* find the sample mean, sample standard deviation, and smaple size
* generate a bootstrap sampling distribution for x-bar using 10,000 resamples
* compute the mean and standard deviation of the bootstrap distribution
* compute a 95% percentile interval to estimate the population mean of repair times
* compute a classical t-based 95% confidence interval 

Are your interval results the same? Is the t-based approach appropriate here? Why or why not? (Note: time is in units of hours.)

So you can check your work:

sample mean: 16.509
sample sd: 19.504
sample size: 23
bootstrap mean: about 16.509
bootstrap sd: about 19.504 / sqrt(23)
bootstrap interval: about (10 to 25)
t-based interval: (8.075 to 24.943)
We need CLT to be true to use t-based approach...

```{r}

```

