---
title: "Normal Distribution Assessment"
subtitle: "In-Class Activity"
author: "PUT YOUR NAME HERE"
date: 'Last rendered on `r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 5
    toc_float: false
---
***

```{r globaloptions, include = FALSE}
# this sets the options but is not shown in the render
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

### Packages Used

```{r packages_used}
library(tidyverse)
library(moments)
```

### Datasets Used

The dataset description can be found [HERE](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/galtonheightdata.html)

```{r}
parent_hts <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/galtonparentheights.csv")
head(parent_hts)
```

### Mother's Height

#### Mean and SD

```{r}
xbar <- mean(parent_hts$Mother)
s <- sd(parent_hts$Mother)
```

#### Density Plot

```{r}
ggplot(parent_hts, aes(x = Mother)) +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(xbar, s),
                color = "orange",
                size = 1)
```

#### Cumulative Distribution Plot

```{r}
ggplot(parent_hts, aes(x = Mother)) +
  stat_ecdf() +
  stat_function(
    fun = pnorm,
    args = list(xbar, s),
    color = "orange",
    size = 1)
```

#### Normal QQ Plot

```{r}
ggplot(parent_hts, aes(sample = Mother)) +
  stat_qq() +
  stat_qq_line(color = "orange",
               size = 1)
```

#### Skewness and Kurtosis

```{r}
skewK <- tibble(
  statistic = c("Skewness", "Kurtosis", "Excess K"),
  theory = c(0, 3, 0),
  data = round(c(skewness(parent_hts$Mother),
                 kurtosis(parent_hts$Mother),
                 kurtosis(parent_hts$Mother) - 3), 2))
print.data.frame(skewK, row.names = FALSE)
```

### Father's Height

#### Density Plot

```{r}

```

#### Cumulative Distribution Plot

```{r}

```

#### Normal QQ Plot

```{r}

```

#### Skewness and Kurtosis

```{r}

```

### Simulations

#### TBA!

```{r}

```

### Related Slides

* [Normal (Gaussian) Distribution: The Math](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/NormalDistributionTheory.html#1)
* [Normal (Gaussian) Distribution: `R` Code](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/NormalDistributionRCode.html#1)
* [Normal (Gaussian) Distribution: Assessment](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/NormalAssessment.html#1)

***
```{r}
sessionInfo()
```
