---
title: "Quantitative EDA"
subtitle: "In-Class Activity"
author: "PUT YOUR NAME HERE"
date: 'Last rendered on `r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 5
    toc_float: false
---
***

```{r globaloptions, include = FALSE}
# this sets the options but is not shown in the render
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

### Packages Used

```{r packages_used}
library(tidyverse)
library(moments)
```

### Datasets Used

#### Bangladesh Wells Dataset
```{r}
wells <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/Bangladesh.csv")
glimpse(wells)
head(wells)
```

#### Happy Face Dataset
```{r}
happyf <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/happyface.csv")
glimpse(happyf)
head(happyf)
```

#### Gas Mileage Dataset
```{r}
mileage <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv")
glimpse(mileage)
head(mileage)
```

#### Old Faithful Dataset
```{r}
geysers <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/oldfaithful.csv")
glimpse(geysers)
head(geysers)
```

### Data Analyses

#### Bangladesh Wells Dataset

##### Plots
```{r}

```

##### Summaries
```{r}

```

#### Happy Face Dataset

##### Plots
```{r}

```

##### Summaries
```{r}

```

#### Gas Mileage Dataset

##### Plots
```{r}

```

##### Summaries
```{r}

```

#### Old Faithful Dataset

##### Plots
```{r}

```

##### Summaries
```{r}

```

Related Slides:

* [EDA Numerical Data: Distribution Shape](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/EDANumericalShape.html#1)
* [EDA Numerical Data: Center, Spread, Z-Scores](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/EDANumericalCSZ.html#1)
* [EDA Numerical Data: Boxplots](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/EDANumericalBox.html#1)

***
```{r}
sessionInfo()
```
