---
title: "Introduction to Sampling"
subtitle: "In-Class Activity"
author: "PUT YOUR NAME HERE"
date: 'Last rendered on `r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 5
    toc_float: false
---
***

```{r globaloptions, include = FALSE}
# this sets the options but is not shown in the render
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

### Packages Used

```{r packages_used}
library(tidyverse)
library(moments)
```

### Uniform Distribution

#### Set Parameters & Visualize

```{r}
a <- 0
b <- 1
```

```{r}
ggplot() + 
  stat_function(fun = dunif, 
                args = list(a, b))
```

#### Sample from the Population

##### Set the Sample Size

```{r}
n1 <- 100000  # sample size -- how many individual values am I choosing from the pop?
```

##### Simulate Data

```{r}
data <- tibble(sample = runif(n1, a, b))

ggplot(data, aes(x = sample)) +
  geom_histogram(bins = 20)
```

##### Assess Normality

```{r}
stats <- data %>% 
  summarize(xbar  = mean(sample),
            s     = sd(sample),
            skew  = skewness(sample),
            ekurt = kurtosis(sample) - 3)
stats

xbar <- stats %>% pull(xbar)
s    <- stats %>% pull(s)

ggplot(data, aes(x = sample)) +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(xbar, s),
                color = "orange",
                size = 1)

ggplot(data, aes(x = sample)) +
  stat_ecdf() +
  stat_function(fun = pnorm,
                args = list(xbar, s),
                color = "orange",
                size = 1)

ggplot(data, aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "orange",
               size = 1)
```

QUESTION: What happens to the distribution of the sample as sample size (n) increases?

ANSWER: As the sample size increases, the sample looks more and more like the parent uniform distribution.

#### Distribution of the Sample Mean

##### Set Sample Size

```{r}
n2 <- 10   # sample size -- how many data points are in each sample?
```

##### Simulate Sample Means

```{r}
N <- 10^5   # choose number of simulations -- how many times are we going to sample?

simresults <- numeric(N)   # initialize storage vector

for (i in 1:N) {
  simresults[i] <- mean(runif(n2, a, b))
}

data <- tibble(sample = simresults)

ggplot(data, aes(x = sample)) +
  geom_histogram(bins = 20) +
  labs(x = "sample means") # my "sample" is 10^5 sample means
```

##### Assess Normality

```{r}
stats <- data %>% 
  summarize(xbar_means  = mean(sample),
            s_means     = sd(sample),
            skew_means  = skewness(sample),
            ekurt_means = kurtosis(sample) - 3)
stats

xbar_means <- stats %>% pull(xbar_means)
s_means    <- stats %>% pull(s_means)

ggplot(data, aes(x = sample)) +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(xbar_means, s_means),
                color = "orange",
                size = 1)

ggplot(data, aes(x = sample)) +
  stat_ecdf() +
  stat_function(fun = pnorm,
                args = list(xbar_means, s_means),
                color = "orange",
                size = 1)

ggplot(data, aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "orange",
               size = 1)
```

QUESTION: What happens to the distribution of the sample *mean* as sample size (n) increases?

ANSWER: As the sample size increases, the sampling distribution of the sample mean becomes bell-shaped (not population shaped).

QUESTION: How large does the sample size have to be before the distribution of the means looks approximately normal? For the uniform distribution, the sampling distribution of the sample mean approaches approximate normality at about n = 10.

ANSWER: 

### Exponential Distribution

#### Set Parameters & Visualize

```{r}
lambda <- 1
```

```{r}
ggplot() + 
  stat_function(fun = dexp, 
                args = list(lambda))
```

#### Sample from the Population

##### Set the Sample Size

```{r}
n1 <- 10000   # sample size
```

##### Simulate Data

```{r}
data <- tibble(sample = rexp(n1, lambda))

ggplot(data, aes(x = sample)) +
  geom_histogram(bins = 20)
```

##### Assess Normality

```{r}
stats <- data %>% 
  summarize(xbar  = mean(sample),
            s     = sd(sample),
            skew  = skewness(sample),
            ekurt = kurtosis(sample) - 3)
stats

xbar <- stats %>% pull(xbar)
s    <- stats %>% pull(s)

ggplot(data, aes(x = sample)) +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(xbar, s),
                color = "orange",
                size = 1)

ggplot(data, aes(x = sample)) +
  stat_ecdf() +
  stat_function(fun = pnorm,
                args = list(xbar, s),
                color = "orange",
                size = 1)

ggplot(data, aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "orange",
               size = 1)
```

QUESTION: What happens to the distribution of the sample as sample size (n) increases?

ANSWER: In the same way the uniform samples converged to the uniform population shape, this exponential samples converge to the exponential shape. In general, samples of values from a population will tend to look more and more like the population as the sample size approaches the population size. They converge to the population.

#### Distribution of the Sample Mean

##### Set Sample Size

```{r}
n2 <- 10000  # sample size
```

##### Simulate Sample Means

```{r}
N <- 10^5   # choose number of simulations

simresults <- numeric(N)   # initialize storage vector

for (i in 1:N) {
  simresults[i] <- mean(rexp(n2, lambda))
}

data <- tibble(sample = simresults)

ggplot(data, aes(x = sample)) +
  geom_histogram(bins = 20)
```

##### Assess Normality

```{r}
stats <- data %>% 
  summarize(xbar_means  = mean(sample),
            s_means     = sd(sample),
            skew_means  = skewness(sample),
            ekurt_means = kurtosis(sample) - 3)
stats

xbar_means <- stats %>% pull(xbar_means)
s_means    <- stats %>% pull(s_means)

ggplot(data, aes(x = sample)) +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(xbar_means, s_means),
                color = "orange",
                size = 1)

ggplot(data, aes(x = sample)) +
  stat_ecdf() +
  stat_function(fun = pnorm,
                args = list(xbar_means, s_means),
                color = "orange",
                size = 1)

ggplot(data, aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "orange",
               size = 1)
```

QUESTION: What happens to the distribution of the sample *mean* as sample size (n) increases?

ANSWER:

QUESTION: If it approaches normality, how large does n have to be to make it approximately normal?

ANSWER:

### Related Slides

* [Normal (Gaussian) Distribution: The Math](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/NormalDistributionTheory.html#1)
* [Normal (Gaussian) Distribution: `R` Code](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/NormalDistributionRCode.html#1)
* [Normal (Gaussian) Distribution: Assessment](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/NormalAssessment.html#1)

***
```{r}
sessionInfo()
```
