---
title: "Introduction to Sampling"
subtitle: "In-Class Activity"
author: "PUT YOUR NAME HERE"
date: 'Last rendered on `r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 5
    toc_float: false
---
***

```{r globaloptions, include = FALSE}
# this sets the options but is not shown in the render
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

### Packages Used

```{r packages_used}
library(tidyverse)
library(moments)
```

### Uniform Distribution

#### Set Parameters & Visualize

```{r}
a <- 0
b <- 1
```

```{r}
ggplot() + 
  stat_function(fun = dunif, 
                args = list(a, b))
```

#### Sample from the Population

##### Set the Sample Size

```{r}
n1 <- 10   # sample size
```

##### Simulate Data

```{r}
data <- tibble(sample = runif(n1, a, b))

ggplot(data, aes(x = sample)) +
  geom_histogram(bins = 20)
```

##### Assess Normality

```{r}
stats <- data %>% 
  summarize(xbar  = mean(sample),
            s     = sd(sample),
            skew  = skewness(sample),
            ekurt = kurtosis(sample) - 3)
stats

xbar <- stats %>% pull(xbar)
s    <- stats %>% pull(s)

ggplot(data, aes(x = sample)) +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(xbar, s),
                color = "orange",
                size = 1)

ggplot(data, aes(x = sample)) +
  stat_ecdf() +
  stat_function(fun = pnorm,
                args = list(xbar, s),
                color = "orange",
                size = 1)

ggplot(data, aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "orange",
               size = 1)
```

QUESTION: What happens to the distribution of the sample as sample size (n) increases?

ANSWER:

#### Distribution of the Sample Mean

##### Set Sample Size

```{r}
n2 <- 1   # sample size
```

##### Simulate Sample Means

```{r}
N <- 10^5   # choose number of simulations

simresults <- numeric(N)   # initialize storage vector

for (i in 1:N) {
  simresults[i] <- mean(runif(n2, a, b))
}

data <- tibble(sample = simresults)

ggplot(data, aes(x = sample)) +
  geom_histogram(bins = 20)
```

##### Assess Normality

```{r}
stats <- data %>% 
  summarize(xbar_means  = mean(sample),
            s_means     = sd(sample),
            skew_means  = skewness(sample),
            ekurt_means = kurtosis(sample) - 3)
stats

xbar_means <- stats %>% pull(xbar_means)
s_means    <- stats %>% pull(s_means)

ggplot(data, aes(x = sample)) +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(xbar_means, s_means),
                color = "orange",
                size = 1)

ggplot(data, aes(x = sample)) +
  stat_ecdf() +
  stat_function(fun = pnorm,
                args = list(xbar_means, s_means),
                color = "orange",
                size = 1)

ggplot(data, aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "orange",
               size = 1)
```

QUESTION: What happens to the distribution of the sample *mean* as sample size (n) increases?

ANSWER:

QUESTION: How large does the sample size have to be before the distribution of the means looks approximately normal?

ANSWER:

### Exponential Distribution

#### Set Parameters & Visualize

```{r}
lambda <- 1
```

```{r}
ggplot() + 
  stat_function(fun = dexp, 
                args = list(lambda))
```

#### Sample from the Population

##### Set the Sample Size

```{r}
n1 <- 10   # sample size
```

##### Simulate Data

```{r}
data <- tibble(sample = rexp(n1, lambda))

ggplot(data, aes(x = sample)) +
  geom_histogram(bins = 20)
```

##### Assess Normality

```{r}
stats <- data %>% 
  summarize(xbar  = mean(sample),
            s     = sd(sample),
            skew  = skewness(sample),
            ekurt = kurtosis(sample) - 3)
stats

xbar <- stats %>% pull(xbar)
s    <- stats %>% pull(s)

ggplot(data, aes(x = sample)) +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(xbar, s),
                color = "orange",
                size = 1)

ggplot(data, aes(x = sample)) +
  stat_ecdf() +
  stat_function(fun = pnorm,
                args = list(xbar, s),
                color = "orange",
                size = 1)

ggplot(data, aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "orange",
               size = 1)
```

QUESTION: What happens to the distribution of the sample as sample size (n) increases?

ANSWER:

#### Distribution of the Sample Mean

##### Set Sample Size

```{r}
n2 <- 1   # sample size
```

##### Simulate Sample Means

```{r}
N <- 10^5   # choose number of simulations

simresults <- numeric(N)   # initialize storage vector

for (i in 1:N) {
  simresults[i] <- mean(rexp(n2, lambda))
}

data <- tibble(sample = simresults)

ggplot(data, aes(x = sample)) +
  geom_histogram(bins = 20)
```

##### Assess Normality

```{r}
stats <- data %>% 
  summarize(xbar_means  = mean(sample),
            s_means     = sd(sample),
            skew_means  = skewness(sample),
            ekurt_means = kurtosis(sample) - 3)
stats

xbar_means <- stats %>% pull(xbar_means)
s_means    <- stats %>% pull(s_means)

ggplot(data, aes(x = sample)) +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(xbar_means, s_means),
                color = "orange",
                size = 1)

ggplot(data, aes(x = sample)) +
  stat_ecdf() +
  stat_function(fun = pnorm,
                args = list(xbar_means, s_means),
                color = "orange",
                size = 1)

ggplot(data, aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "orange",
               size = 1)
```

QUESTION: What happens to the distribution of the sample *mean* as sample size (n) increases?

ANSWER:

QUESTION: How large does the sample size have to be before the distribution of the means looks approximately normal?

ANSWER:

### Related Slides

* [Normal (Gaussian) Distribution: The Math](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/NormalDistributionTheory.html#1)
* [Normal (Gaussian) Distribution: `R` Code](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/NormalDistributionRCode.html#1)
* [Normal (Gaussian) Distribution: Assessment](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/NormalAssessment.html#1)

***
```{r}
sessionInfo()
```
