---
title: "Think-Pair-Share: Creating Vectors"
author: "SOLUTIONS // Jill E. Thomley"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

***
```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

**Original Response Post:** Try to create the following vectors in the most efficient way possible and post your code, your best attempt, or some comment indicating where you got stuck. Each can be created with a single line of code using c(), colon, seq(), rep(), sample(), or some combination.  

1) All the possible readings of a machine that measures pH to one decimal of precision.

```{r}
(pH <- seq(from = 0, to = 14, by = .1))
```

**Note:** As you get comfortable with the `seq` function, you can omit the parameter names and write it as `seq(0, 14, .1)`. Putting `()` around the created object prints it out as well.

2) A countdown clock counting backwards from 1 minute to zero in whole-second intervals.

```{r}
clock <- 60:0
print(clock)
```

**Note:** Many people created this vector using `c(60:0)`, but the `c()` is not needed when making a sequence with a colon. We also do not need to use the `seq()` function because we are incrementing by 1. One creative person formatted the vector to look like a countdown clock. This is interesting, but the output would be a character vector versus numeric, so it depends on the purpose for making the vector. Nice thinking with `sprintf()`!

For more on the `paste()` and `sprintf()` functions, see http://www.cookbook-r.com/Strings/Creating_strings_from_variables/. They are particularly useful when embedding output as inline code in R Markdown.

```{r}
c("1:00", sprintf("%i:%2.2i", 0, 59:0))
```

3) A particular song's chorus contains the following three phrases, each sung once then repeated three more times in order: "I get knocked down", "But I get up again", "You're never gonna keep me down".

```{r}
(song <- rep(c("I get knocked down", "But I get up again", "You're never gonna keep me down"), times = 4))
```

**Note:** The chorus is sung once and repeated three *more* times, so the total number of repetitions is 4.

4) In the US population, there are four blood types distributed as follows: A = 40%, B = 11%, AB = 4%, and O = 45%. Create a vector that represents the blood types of 100 people in those exact ratios (i.e., your vector should contain 100 values A, B, AB, O in the exact proportions indicated).

```{r}
(bloodtypes <- rep(c("A", "B", "AB", "O"), times = c(40, 11, 4, 45)))
```

**Note:** Multiplying the percentages by 100 tells us how many times each blood type needs to be repeated. Many people created the vector using multiple `rep()` functions, such as `c(rep("AB", 4), rep("A", 40), rep("B", 11), rep("O", 45))`. This produces the same results, but as we can see, the `rep()` function already allows for non-equal numbers of repetitions.

5) The blood types of a random sample of 100 people from the US population, given the population distribution previously indicated. 

```{r}
bloodsample1 <- sample(c("A", "B", "AB", "O"), size = 100, replace = TRUE, prob = c(.4, .11, .04, .45))

bloodsample2 <- sample(bloodtypes, size = 100, replace = TRUE)
```

**Note:** The first solution above is the better one, since it is self-contained. It is important to specify `replace = TRUE`, otherwise you get a reshuffling (permutation) of the values rather than a random sample. The second solution shows how we can sample from a vector that has already been created, which we will do with sample data in the future. Since the blood types exist in the vector in the correct proportions, the relative *probability* of selecting each of the four types is preserved. The examples below show three random permutations versus three random samples.

```{r}
(permutation1 <- sample(bloodtypes, size = 100))
table(permutation1)
(permutation2 <- sample(bloodtypes, size = 100))
table(permutation2)
(permutation3 <- sample(bloodtypes, size = 100))
table(permutation3)
(randsample1 <- sample(bloodtypes, size = 100, replace = TRUE))
table(randsample1)
(randsample2 <- sample(bloodtypes, size = 100, replace = TRUE))
table(randsample2)
(randsample3 <- sample(bloodtypes, size = 100, replace = TRUE))
table(randsample3)
```


***
```{r}
sessionInfo()
```

