---
title: "MSRR CH02, Sections 4-6 HW"
author: "SOLUTIONS / Jill E. Thomley"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = "")
```

***

**Instructions**

Complete the following problems from _Mathematical Statistics with Resampling and R_. You can access the textbook datasets on our [Class Data Files](https://stat-jet-asu.github.io/Datasets/DatasetList.html) page or from the `R` package `resampledata`. The textbook examples use base `R`code for plotting, but you `ggplot2` for visualizations. You may use `table()` to make tables, but you should use `dplyr` for other summaries. Refer to the case studies in Chapter 1 to add informative titles and axis labels to your plots and to provide context for your answers.

Load the packages and datasets you need here.

```{r}

```


***
#### Problem 2.7

Part (a)

Note: For summaries, include sample size, mean, standard deviation, five-number summary, skewness, and kurtosis.

```{r}

```

Part (b)

Note: Also make a density plot and ecdf plot. As we learned to do in class, include theoretical normal reference curves.

```{r}

```

ANSWER: 

Part (c)

```{r}

```

Part (d)

Note: For summaries, include sample size, mean, standard deviation, five-number summary, skewness, and kurtosis. Use `dplyr` functions rather than `tapply()`.

```{r}

```

Part (e)

```{r}

```

ANSWER: 


***
#### Problem 2.8

Note: Where the problem asks you to "describe the distribution," discuss the shape and possible outliers. Choose an appropriate number of bins for the histograms.

Part (a)

```{r}

```

ANSWER: 

Part (b)

Note: Use `mutate` and name the new variable `m.cpc_diff`.

```{r}

```

ANSWER: 

Part (c)

```{r}

```

ANSWER: 


***

```{r}
sessionInfo()
```

