---
title: "MSRR CH02, Sections 4-6 HW"
author: "SOLUTIONS / Jill E. Thomley"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = "")
```

***

**Instructions**

Complete the following problems from _Mathematical Statistics with Resampling and R_ and additional example problems using the normal distribution functions in `R`. You can access the textbook datasets on our [Class Data Files](https://stat-jet-asu.github.io/Datasets/DatasetList.html) page or from the `R` package `resampledata`. The textbook examples use base `R`code for plotting, but you `ggplot2` for visualizations. You may use `table()` to make tables, but you should use `dplyr` for other summaries. Refer to the case studies in Chapter 1 to add informative titles and axis labels to your plots and to provide context for your answers.

Load the packages and datasets you need here.

```{r}

```


***
#### Problem 2.7

Part (a)

Note: For summaries, include sample size, mean, standard deviation, five-number summary, skewness, and kurtosis.

```{r}

```

Part (b)

Note: Also make a density plot and ecdf plot. As we learned to do in class, include theoretical normal reference curves.

```{r}

```

ANSWER: 

Part (c)

```{r}

```

Part (d)

Note: For summaries, include sample size, mean, standard deviation, five-number summary, skewness, and kurtosis. Use `dplyr` functions rather than `tapply()`.

```{r}

```

Part (e)

```{r}

```

ANSWER: 


***
#### Problem 2.8

Note: Where the problem asks you to "describe the distribution," discuss the shape and possible outliers. Choose an appropriate number of bins for the histograms.

Part (a)

```{r}

```

ANSWER: 

Part (b)

Note: Use `mutate` and name the new variable `m.cpc_diff`.

```{r}

```

ANSWER: 

Part (c)

```{r}

```

ANSWER: 


***
#### `R` Problem 1

Let the random variable X have a standard normal distribution.

Part (a) : What is P(X < 1.5)?

```{r}

```

Part (b) : What is P(-2.0 < X < 1.5)?

```{r}

```

Part (c) : What is P(X < -1.75 or X > 1.75)?

```{r}

```

Part (d) : What value of x satisfies P(X < x) = 0.45?

```{r}

```

Part (e) : What value of x satisfies P(X ≥ x) = 0.35?

```{r}

```

Part (f) : The middle 50% of the area under a standard normal curve lies between what values?

```{r}

```


***
#### `R` Problem 2

In Hottsylvania, the mean July midday temperature is 32 °C with a standard deviation of 3 °C. Assume that midday temperatures can be modeled with an approximately normal distribution. July has 31 days.

Part (a) : About how many days should be less than 29.44 °C (about 85 °F)?

```{r}

```

Part (b) : How many days should have a midday temperature between 35 °C and 38 °C?

```{r}

```

Part (c) : The midday temperature of the hottest 30% of days should be greater than what temperature?

```{r}

```


***
#### `R` Problem 3

Generate a random sample of size 100 from a standard normal distribution. Create a density plot and a normal probability plot with appropriate theoretical reference curves.

```{r}

```


***

```{r}
sessionInfo()
```

