---
title: "MSRR CH02, Sections 1-3 HW"
author: "YOUR NAME HERE"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = "")
```

***

**Instructions**

Complete the following problems from _Mathematical Statistics with Resampling and R_. You can access the textbook datasets on our [Class Data Files](https://stat-jet-asu.github.io/Datasets/DatasetList.html) page or from the `R` package `resampledata`. The textbook examples use base `R`code for plotting, but you `ggplot2` for visualizations. You may use `table()` to make tables, but you should use `dplyr` for other summaries. Refer to the case studies in Chapter 1 to add informative titles and axis labels to your plots and to provide context for your answers.

Load the packages and datasets you need here.

```{r}

```


***
#### Problem 2.1

Note: Test for equality using code, not just visual inspection.

```{r}

```


***
#### Problem 2.2

Note: Test for equality using code, not just visual inspection.

```{r}

```


***
#### Problem 2.5

Note: In (c), create the table of proportions you need to answer (d) as well as a table of counts.

Part (a)

```{r}

```

Part (b)

```{r}

```

ANSWER: The summary shows the NA values while the table does not.

Part (c)

```{r}

```

Part (d)

ANSWER: 


***
#### Problem 2.6

Note: In (b), create a table of proportions as well as a table of counts to answer the question. In (c), also compute the upper fence for each group to show the upper cutoff for outliers.

Part (a)

```{r}

```

Part (b)

```{r}

```

Part (c)

```{r}

```

Comparison 1: 

Comparison 2: 

Comparison 3: 

Part (d)

```{r}

```

Part (e)

```{r}

```

ANSWER:


***

```{r}
sessionInfo()
```

