---
title: "MSRR CH09, Sections 1-3 HW"
author: "YOUR NAME HERE"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = "")
```

***

**Instructions**

Complete the following problems from _Mathematical Statistics with Resampling and R_. You can access the textbook datasets on our [Class Data Files](https://stat-jet-asu.github.io/Datasets/DatasetList.html) page or from the `R` package `resampledata`. The textbook examples use base `R`code for plotting, but you `ggplot2` for visualizations. You may use `table()` to make tables, but you should use `dplyr` for other summaries. Refer to the case studies in Chapter 1 to add informative titles and axis labels to your plots and to provide context for your answers.

Load the packages and datasets you need here.

```{r}

```


***
#### Problem 9.6

Part (a) 

```{r}

```

Part (b)

Note: For "What do you observe?", comment on linearity, direction of trend, and outliers.

```{r}

```

ANSWER: 

Part (c)

Hint: Use `filter()` with `Weight` to remove the outliers.

```{r}

```

ANSWER: 

```{r}

```


***
#### Problem 9.8

Part (a)

```{r}

```

Part (b)

```{r}

```

ANSWER: 


***
#### Problem 9.9

Part (a)

Note: Also make a scatterplot of field goal percentage versus offensive rebounds.

```{r}

```

Part (b)

```{r}

```

Part (c)

```{r}

```


***
#### Problem 9.16

Note: In (b), create a table of proportions as well as a table of counts to answer the question. In (c), also compute the upper fence for each group to show the upper cutoff for outliers.

Part (a)

```{r}

```

Part (b)

```{r}

```

INTERPRETATION:

Part (c)

```{r}

```

INTERPRETATION:

Part (d) ADDED Perform a residual analysis like [MD 6.1.4](https://moderndive.com/6-regression.html#model1) and comment on the appropriateness of a straight-line model. 

```{r}

```

ANSWER:


***

```{r}
sessionInfo()
```

