---
title: "Bernoulli and Binomial Distributions: `R` Code"
author: "Author: Jill E. Thomley"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  ioslides_presentation:
    logo: images/logoASU.jpg
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(comment="", message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(knitr)
```

## Before We Begin...

These slides are not meant to be standalone information. You should take notes to flesh out the contents. I recommend that you create an R Markdown document where you can combine information and code from the slides and your own additional notes and explorations to make connections.

**Related Materials**

* Appendix A of *Mathematical Statistics with Resampling and R, 2^nd^ Ed.* (Review of Probability)
* DataCamp [Foundations of Probability in R](https://learn.datacamp.com/courses/foundations-of-probability-in-r)
* [Bernoulli and Binomial Distributions: The Math](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/BinomialDistributionTheory.html#1) class slides

Refer to "[
4.1 ioslides presentation](https://bookdown.org/yihui/rmarkdown/ioslides-presentation.html)" in [*R Markdown: The Definitive Guide*](https://bookdown.org/yihui/rmarkdown/) if you are interested in creating R Markdown slides.


## The Binomial Model

$$X \sim Bin(n,p)$$

    x = number of successes = {0, 1, 2, ..., n}
    n = number of trials
    p = probability of success on a single trial
    1 - p = probability of failure on a single trial

$$f(x) = P(X = x) = {n \choose x}p^x(1-p)^{n-x}$$

$$F(x)=P(X \leq x) = \sum\limits_{i=0}^x f(x) = \sum\limits_{i=0}^x {n \choose i}p^i(1-p)^{n-i}$$

$$E[X] = np, Var[X] = np(1-p)$$

## Calculating the pmf f(x)

The `dbinom()` function is used to calculate pmf probabilities for the binomial distribution.

$P(X = x)$ = $f(x,n,p)$ = `dbinom(x, size, prob)` 

Consider a biased coin example where $X \sim Bin(3, 0.7)$.

```{r}
dbinom(1, 3, 0.7)   # P(X = 1 head)
dbinom(3, 3, 0.7)   # P(X = 3 heads)
```


## Vectorized Calculations

The `dbinom` function can accept vectors as input. It loops over all values in the input vector and outputs a vector the same length. What are the probabilities for $x = \{0, 1, 2, 3\}$ heads if we have a fair coin? We can compute the entire pdf at once.

Recall that in `R` we can easily make a vector of integers using the `:` operator, without needing to use `c()`.

```{r}
dbinom(0:3, 3, 0.5)
```

Notice that when $p = 0.5$, the distribution is symmetric. Refer to $f(x)$ and see if you can figure out why that is true.


## Calculating the CDF F(x)

The `pbinom()` function is used to calculate CDF probabilities for the binomial distribution. It also accepts vector inputs.

$P(X \leq x)$ = $F(x,n,p)$ = `pbinom(q, size, prob)`

```{r}
pbinom(0:3, 3, 0.7)   # biased coin, x = {0, 1, 2, 3} heads
```

```{r}
pbinom(0:3, 3, 0.5)   # fair coin, x = {0, 1, 2, 3} heads
```

The CDF for the fair coin also has a kind of symmetry to it. Why is that the case?


## The Upper Tail of the Distribution

```{r}
pbinom(2, 3, 0.7)    # probability of 0, 1, or 2 heads
1 - pbinom(2, 3, 0.7)   # probability of > 2 = 3 heads
pbinom(2, 3, 0.7, lower.tail = FALSE)   # better code
```

Pay close attention to $<$ vs. $\leq$ and $>$ vs. $\geq$ with `pbinom()`. We are dealing with discrete values in the binomial distribution.


## Lower and Upper Tails Relationship

For three tosses of our fair coin...

```{r}
pbinom(0:3, 3, 0.5) # P(X <= x) for all possible x
```

```{r}
pbinom(0:3, 3, 0.5, lower.tail = FALSE) # P(X > x) for all possible x
```

```{r}
pbinom(0:3, 3, 0.5) + pbinom(0:3, 3, 0.5, lower.tail = FALSE)
```

The two halves sum to 1. Voila! Can you picture it on the plot?


##

Can you show on the plot the difference between `lower.tail = TRUE` and `lower.tail = FALSE` for a given $x$?

```{r, echo = FALSE}
distTable <- tibble(x = 0:3, 
                    pdf = dbinom(0:3, 3, 0.7), 
                    CDF = pbinom(0:3, 3, 0.7)) 
ggplot(distTable, aes(x = x, y = pdf, label = pdf)) +
  geom_col(fill = "skyblue", color = "black", width = 0.75) +
  scale_y_continuous(limits = c(0, 1.05),
                     breaks = seq(0, 1, 0.05),
                     expand = c(0,0)) +
  geom_label(vjust = 0, nudge_y = 0.05) +
  labs(title = "Probability Mass Function of X ~ Bin(3, 0.7)",
       x = "number of successes",
       y = "probability") +
  theme_linedraw()
```


## Calculating the Inverse of the CDF

Until now, we have been finding probabilities associated with given values of $x$. What if we have the probability and need to find $x$? For a given quantile (percentile), the inverse CDF gives the value of $x$ that makes $P(X \leq x) = quantile$ true.

$q^{th}$ percentile = $F^{-1}(x,n,p)$ = `qbinom(p, size, prob)` 

```{r}
qbinom(0.25, 3, 0.7)   # 25th percentile of X ~ Bin(3, 0.7)
qbinom(0.75, 3, 0.7)   # 75th percentile of X ~ Bin(3, 0.7)
```


## Generating Random Values

To generate a (*pseudo*)random sample from a given binomial distribution, use `rbinom (n, size, prob)`. The binomial $n$ parameter is called *size* here, since $n$ is standard notation for sample size in statistics. 

Simulate 20 random values from biased coin $X \sim Bin(3, 0.7)$

```{r}
rbinom(20, 3, 0.7)   
```

Each output value in this random sample is $x$ = *number of heads* (successes) in a sequence of 3 flips. We have 20 replicates of the coin flipping "experiment".


## Summarizing Random Samples

```{r}
binom_sample <- tibble(x = rbinom(10^5, 3, 0.7))   # 100,000 samples

binom_sample %>% 
  summarize(mean_x = mean(x),
            var_x  = var(x))
```

As the sample size approaches $\infty$, sample mean $\bar{x}$ and sample variance $s^2$ converge to E[X] and Var[X]. Because of this, we can use simulation / random sampling to *estimate* parameters when the underlying distribution is not known. 


## Creating a Distribution "Table"

```{r}
n <- 3     # enter n
p <- 0.7   # enter p
distTable <- tibble(x = 0:n, 
                    pmf = dbinom(0:n, n, p), 
                    CDF = pbinom(0:n, n, p)) 
distTable
```

You can format for digits and appearance with `kable()`.


## Creating a Plot of the pmf

The plot uses the distribution table we created.

```{r, eval = FALSE}
breakby <- 1   # choose spacing for the x-axis tick marks

ggplot(distTable, aes(x = x, y = pmf)) +
  geom_col(fill = "skyblue", color = "black") +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "f(x)")
```

Add additional formatting or themes as desired.

To display more decimal places for $p$ in the plot title, change %1.1f to %1.*decimalplaces*f in the code (e.g., %1.3f for three).


## The Output

```{r, echo = FALSE}
breakby <- 1   # choose spacing for the x-axis tick marks

ggplot(distTable, aes(x = x, y = pmf)) +
  geom_col(fill = "skyblue", color = "black") +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "f(x)")
```


## Creating a Plot of the CDF

The plot uses the distribution table we created.

```{r, eval = FALSE}
breakby <- 1   # choose spacing for the x-axis tick marks

ggplot(distTable, aes(x = x, y = CDF)) +
  geom_col(fill = "skyblue", color = "black") +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "F(x)")
```

Add additional formatting or themes as desired.

To display more decimal places for $p$ in the plot title, change %1.1f to %1.*decimalplaces*f in the code (e.g., %1.3f for three).


## The Output

```{r, echo = FALSE}
breakby <- 1   # choose spacing for the x-axis tick marks

ggplot(distTable, aes(x = x, y = CDF)) +
  geom_col(fill = "skyblue", color = "black") +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "F(x)")
```


## Example for 20 Fair Coin Flips

```{r}
n <- 20    # enter n
p <- 0.5   # enter p
distTable <- tibble(x = 0:n, 
                    pmf = dbinom(0:n, n, p), 
                    CDF = pbinom(0:n, n, p)) 
head(distTable, 4)
```


## The pmf Plot

```{r, echo = FALSE}
breakby <- 2   # choose spacing for the x-axis tick marks

ggplot(distTable, aes(x = x, y = pmf)) +
  geom_col(fill = "skyblue", color = "black") +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "f(x)")
```


## The CDF Plot

```{r, echo = FALSE}
breakby <- 2   # choose spacing for the x-axis tick marks

ggplot(distTable, aes(x = x, y = CDF)) +
  geom_col(fill = "skyblue", color = "black") +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "F(x)")
```