---
title: "Bernoulli and Binomial Distributions: `R` Code"
author: "Author: Jill E. Thomley"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  ioslides_presentation:
    logo: images/logoASU.jpg
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
library(tidyverse)
```

## Before We Begin...

These slides are not meant to be standalone information. You should take notes to flesh out the contents. I recommend that you create an R Markdown document where you can combine information and code from the slides and your own additional notes and explorations to make connections.

**Related Materials**

* Appendix A of *Mathematical Statistics with Resampling and R, 2^nd^ Ed.* (Review of Probability)
* DataCamp [Foundations of Probability in R](https://learn.datacamp.com/courses/foundations-of-probability-in-r)
* [Bernoulli and Binomial Distributions: The Math](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/BinomialDistributionTheory.html) class slides


## Index

* Slides 02-03 &mdash; Materials & Index
* Slides 04-13 &mdash; Calculating PMF and CDF
   + Slides 05-07 &mdash; General Model and PMF
   + Slides 08-13 &mdash; CDF and Inverse CDF
* Slides 14-19 &mdash; Simulating a Binomial
   + Slides 15-17 &mdash; Generating and Tabulating
   + Slides 18-19 &mdash; Plotting and Summarizing
* Slides 20-28 &mdash; Creating a Distribution “Table” and Plots
   + Slides 21-25 &mdash; General Algorithm and `R` Code
   + Slides 26-28 &mdash; Example for 20 Fair Coin Flips


# Calculating PMF and CDF

## The Binomial Model

$$X \sim Bin(n,p)$$

    x = number of successes = {0, 1, 2, ..., n}
    n = number of trials
    p = probability of success on a single trial
    1 - p = probability of failure on a single trial

$$f(x) = P(X = x) = {n \choose x}p^x(1-p)^{n-x}$$

$$F(x)=P(X \leq x) = \sum\limits_{i=0}^x f(x) = \sum\limits_{i=0}^x {n \choose i}p^i(1-p)^{n-i}$$

$$E[X] = np \text{ and }Var[X] = np(1-p)$$

## Calculating the pmf or f(x)

The `dbinom()` function is used to calculate pmf probabilities for the binomial distribution.

$P(X = x)$ = $f(x,n,p) \rightarrow$ `dbinom(x, size, prob)` in `R`

Consider our biased coin example where $X \sim Bin(3, 0.7)$.

```{r}
dbinom(1, 3, 0.7)   # P(X = 1 head)
dbinom(3, 3, 0.7)   # P(X = 3 heads)
```


## Vectorized Calculations

The `dbinom` function can accept vectors as input. It loops over all values in the input vector and outputs a vector the same length. What are the probabilities for $x = \{0, 1, 2, 3\}$ heads if we have a fair coin? We can compute the entire pdf at once.

Recall that in `R` we can easily make a vector of integers using the `:` operator, without needing to use `c()`.

```{r}
dbinom(0:3, 3, 0.5)   # compute all probabilities at the same time
```

Notice that when $p = 0.5$, the distribution is symmetric. Refer to $f(x)$ and see if you can figure out why that is true.


## Calculating the CDF or F(x)

The `pbinom()` function is used to calculate CDF probabilities for the binomial distribution. It also accepts vector inputs.

$P(X \leq x)$ = $F(x,n,p) \rightarrow$ `pbinom(q, size, prob)` in `R`

```{r}
pbinom(0:3, 3, 0.7)   # biased coin, x = {0, 1, 2, 3} heads
```

```{r}
pbinom(0:3, 3, 0.5)   # fair coin, x = {0, 1, 2, 3} heads
```

The CDF for the fair coin also has a kind of symmetry to it. Why is that the case?


## The Upper Tail of the Distribution

```{r}
pbinom(2, 3, 0.7)    # probability of 0, 1, or 2 heads
1 - pbinom(2, 3, 0.7)   # probability of more than 2 heads
pbinom(2, 3, 0.7, lower.tail = FALSE)   # better code
```

Pay close attention to $<$ vs $\leq$ and $>$ vs $\geq$ with `pbinom()`. We are dealing with discrete sample spaces here.


##

For three tosses of our fair coin...

```{r}
pbinom(0:3, 3, 0.5)   # P(X <= x) for all possible x
```

```{r}
pbinom(0:3, 3, 0.5, lower.tail = FALSE)   # P(X > x)
```

```{r}
pbinom(0:3, 3, 0.5) + pbinom(0:3, 3, 0.5, lower.tail = FALSE)
```

The two halves sum to 1. Voila! Can you picture it on a graph?


##

Can you show on the pmf graph below the difference between `lower.tail = TRUE` and `lower.tail = FALSE` for a given $x$?

```{r, echo = FALSE}
distTable <- tibble(x = 0:3, 
                    pdf = dbinom(0:3, 3, 0.7), 
                    CDF = pbinom(0:3, 3, 0.7)) 
ggplot(distTable, aes(x = x, y = pdf, label = pdf)) +
  geom_col(fill = "skyblue", color = "black", width = 0.75) +
  scale_y_continuous(limits = c(0, 1.05),
                     breaks = seq(0, 1, 0.05),
                     expand = c(0,0)) +
  geom_label(vjust = 0, nudge_y = 0.05) +
  labs(title = "Probability Mass Function of X ~ Bin(3, 0.7)",
       x = "x = number of successes",
       y = "dbinom(x, n, p) = probability") +
  theme_linedraw()
```


## 

The areas included in the calculation are shaded in solid black.

```{r, echo = FALSE, fig.height = 5}
plotA <- ggplot(distTable, aes(x = x, y = pdf, label = pdf)) +
  geom_col(fill = c("black", "black", "black", "skyblue"), color = "black", width = 0.75) +
  scale_y_continuous(limits = c(0, 1.05),
                     breaks = seq(0, 1, 0.1),
                     expand = c(0,0)) +
  geom_label(vjust = 0, nudge_y = 0.05) +
  labs(title = sprintf("P(X ≤ 2) = pbinom(2, 3, 0.7) = %1.3f", pbinom(2, 3, 0.7)),
       x = "x = number of successes",
       y = "dbinom(x, n, p) = probability") +
  theme_linedraw()

plotB <- ggplot(distTable, aes(x = x, y = pdf, label = pdf)) +
  geom_col(fill = c("skyblue", "skyblue", "skyblue", "black"), color = "black", width = 0.75) +
  scale_y_continuous(limits = c(0, 1.05),
                     breaks = seq(0, 1, 0.1),
                     expand = c(0,0)) +
  geom_label(vjust = 0, nudge_y = 0.05) +
  labs(title = sprintf("P(X > 2) = pbinom(2, 3, 0.7, lower.tail = FALSE) = %1.3f", pbinom(2, 3, 0.7, F)),
       x = "x = number of successes",
       y = "dbinom(x, n, p) = probability") +
  theme_linedraw()

gridExtra::grid.arrange(plotA, plotB)
```


## Calculating the Inverse of the CDF

Until now, we have been finding probabilities associated with given values of $x$. What if we have the probability and need $x$? For a given quantile (percentile), the inverse CDF gives us the $x$ that makes $P(X \leq x) = quantile$ (approximately) true.

$q^{th}$ percentile = $F^{-1}(x,n,p) \rightarrow$ `qbinom(p, size, prob)` in `R`

```{r}
qbinom(0.25, 3, 0.7)   # 25th percentile of X ~ Bin(3, 0.7)
qbinom(0.75, 3, 0.7)   # 75th percentile of X ~ Bin(3, 0.7)
```


# Simulating a Binomial

## Generating Random Values

To generate a (*pseudo*)random sample from a given binomial distribution, use `rbinom(n, size, prob)`. The binomial $n$ parameter is called *size* in `R`'s functions, because $n$ is standard notation for sample size in statistics. 

Simulate 20 random values from biased coin $X \sim Bin(3, 0.7)$

```{r}
rbinom(20, 3, 0.7)   
```

Each output value in this random sample is $x$ = *number of heads* (successes) in a sequence of 3 flips. We have 20 replicates of the "experiment" of flipping 3 coins.


## Tabulating Random Samples

For a discrete random variable with a small number of possible values, we can tabulate the simulated values. How close are the simulated proportions to the theoretical probabilities?

```{r}
binom_sample <- tibble(x = rbinom(10^5, 3, 0.7))   # 100,000 samples
binom_sample %>% 
  count(x) %>% 
  mutate(proportion = prop.table(n))
```


## Empirical vs Theoretical

Values obtained from experimentation, including simulation, are *empirical* values. Empirical proportions can be used to estimate unknown probabilities.

```{r}
binom_sample %>% 
  count(x) %>% 
  mutate(empirical   = prop.table(n),
         theoretical = dbinom(0:3, 3, 0.7))
```


## Plotting Random Samples

For a discrete random variable with a small number of values, we can create a bar plot. More often, we will use a histogram to accommodate a larger range of possible x values.

```{r, fig.height = 3}
ggplot(binom_sample, aes(x = x)) + geom_bar()   # add formatting 
```


## Summarizing Random Samples

How close are these values to the theoretical E[X] and Var[X]?

```{r}
binom_sample %>% 
  summarize(samples = n(),
            mean_x  = mean(x),
            var_x   = var(x))
```

As the sample size approaches $\infty$, sample mean $\bar{x}$ and sample variance $s^2$ converge to E[X] and Var[X]. Because of this, we can use simulation / random sampling to *estimate* parameters when the underlying distribution is not known. 


# Distribution “Table” and Plots of the PMF and CDF

## Creating a Distribution "Table"

```{r}
n <- 3     # enter n
p <- 0.7   # enter p
distTable <- tibble(x   = 0:n,               # x = {0, 1, ... , n}
                    pmf = dbinom(x, n, p),   # calculate for all x
                    CDF = pbinom(x, n, p))   # calculate for all x
distTable
```

You can format for digits and appearance with [`kable()`](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html).


## Creating a Plot of the PMF

The plot uses the $n$ and $p$ we input and `distTable` object.

```{r, eval = FALSE}
breakby <- 1   # choose spacing for the x-axis tick marks

ggplot(distTable, aes(x = x, y = pmf)) +
  geom_col(fill = "skyblue", color = "black", width = 1) +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("PMF of X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "f(x)") +
  theme_linedraw()
```

You can add other formatting or change themes as desired.

To display more decimal places for $p$ in the plot title, change %1.1f to %1.*decimalplaces*f in the code (e.g., %1.3f for three).


## The Output

```{r, echo = FALSE}
breakby <- 1   # choose spacing for the x-axis tick marks

ggplot(distTable, aes(x = x, y = pmf)) +
  geom_col(fill = "skyblue", color = "black", width = 1) +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("PMF of X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "f(x)") +
  theme_linedraw()
```


## Creating a Plot of the CDF

The plot uses the $n$ and $p$ we input and `distTable` object.

```{r, eval = FALSE}
breakby <- 1   # choose spacing for the x-axis tick marks

ggplot(distTable, aes(x = x, y = CDF)) +
  geom_col(fill = "skyblue", color = "black", width = 1) +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("CFD of X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "F(x)") +
  theme_linedraw()
```

Add additional formatting or themes as desired.

To display more decimal places for $p$ in the plot title, change %1.1f to %1.*decimalplaces*f in the code (e.g., %1.3f for three).


## The Output

```{r, echo = FALSE}
breakby <- 1   # choose spacing for the x-axis tick marks

ggplot(distTable, aes(x = x, y = CDF)) +
  geom_col(fill = "skyblue", color = "black", width = 1) +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("CDF of X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "F(x)") +
  theme_linedraw()
```


## Example for 20 Fair Coin Flips

```{r}
n <- 20
p <- 0.5
distTable <- tibble(x   = 0:n,
                    pmf = dbinom(x, n, p),
                    CDF = pbinom(x, n, p))
head(distTable, 6)   # the full table is too large for the slide
```


## The pmf Plot

```{r, echo = FALSE}
breakby <- 2

ggplot(distTable, aes(x = x, y = pmf)) +
  geom_col(fill = "skyblue", color = "black", width = 1) +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("PMF of X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "f(x)") +
  theme_linedraw()
```

For the future: What shape does this (approximately) look like?


## The CDF Plot

```{r, echo = FALSE}
breakby <- 2

ggplot(distTable, aes(x = x, y = CDF)) +
  geom_col(fill = "skyblue", color = "black", width = 1) +
  scale_x_continuous(breaks = seq(from = 0, to = n, by = breakby)) +
  labs(title = sprintf("CDF of X ~ Bin(%1.0f, %1.1f)", n, p),
       x = "x", 
       y = "F(x)") +
  theme_linedraw()
```

For the future: Notice the symmetric logistic (S-curve) shape.
