---
title: "EDA &#9654; Categorical Data<br>Factors"
author: "Author: Jill E. Thomley"
date: '`r format(Sys.time(), "%B %d, %Y @ %I:%M %p")`'
output: 
  ioslides_presentation:
    logo: images/logoASU.jpg
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
library(tidyverse)
```


## Before We Begin...

These slides are not meant to be standalone information. You should take notes to flesh out the contents. I recommend that you create an R Markdown document where you can combine information and code from the slides and your own additional notes and explorations to make connections.

**Related Materials**

* CH14 of [R for Data Science](https://r4ds.had.co.nz/strings.html)
* CH15 of [R for Data Science](https://r4ds.had.co.nz/factors.html)
* [Understanding Factors](http://monashbioinformaticsplatform.github.io/2015-09-28-rbioinformatics-intro-r/01-supp-factors.html) from the [Software Carpentry](https://software-carpentry.org/) R course.


## What is a Factor in `R`?

By default, `R` displays the possible values of a character vector in ascending alphabetical order. When we summarize, visualize, or model with categorical variables, we often want their categories to appear in a different order. 

Structurally, factor variables are stored as unique integers with attached labels that are used for display. This is similar to how `logical` vectors display `TRUE` and `FALSE` but are stored as 0's and 1's "under the hood". 

The set of possible values of a factor are are known as levels.

The `factor` function is used to create and modify factors in base `R` and the `tidyverse` contains a package called `forcats` that offers other useful functions.


## Revisiting An Example

The [STT3850majors.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/stt3850majors.csv) dataset contains information regarding students in Dr. Thomley's past STT3850 classes. 

`students <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/stt3850majors.csv"))`

```{r echo=FALSE}
students <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/stt3850majors.csv"))
```

* College = ASU college the student is in (e.g., Arts & sciences)
* Year = student's class year (e.g., freshman, sophomore)
* BannerMajor = student's major as listed in banner
* CodedMajor = Dr. Thomley's recoded list of majors
* Semester = the semester when the student took STT3850
* AY = academic year (e.g., 2018-2019)


## Dataset Structure

```{r}
str(students) # note that all variables are character variables
```


## Class Years

Notice the `Year` categories are not in the standard order. We should probably convert the `character` variable to a `factor`. Then we can choose an order other than alphabetical.

```{r}
students %>% 
  count(Year) %>% 
  mutate(percents = 100 * prop.table(n))
```


## Converting to a Factor

We know from our previous table that there are five categories or levels. Let's create a vector that puts all of the levels in the order we want them. Then create our factor using `factor()` to replace the original version.

```{r}
Year_levels <- c("Freshman", "Sophomore", "Junior", "Senior", 
                 "Graduate Certificate Program")

students$Year <- factor(students$Year, levels = Year_levels)

str(students$Year)
```


## Now the `tidyverse` Way

We can also create a factor using the `mutate()` function. Here we save the new version of `Year` over the original version rather than creating a new variable. We also re-store the transformed dataset under the same name `students`. You can use `str()` on the dataset to see how it is different than before.

```{r}
students <- students %>% 
  mutate(Year = factor(Year, levels = Year_levels))

levels(students$Year)
```


## Dataset Structure

```{r}
str(students) # note that Year is now a factor variable
```


## Converting Several Variables

The advantage of `tidyverse` way is being able to recode many variables at once. You can also embed the levels vectors rather than creating them separately. For presentation spacing, I broke the levels vectors across lines. This is OK in `R` since the program is looking for the closing `)` as the end-of-function.

```{r}
students <- students %>% 
  mutate(Year = factor(Year, levels = c("Freshman",
                                        "Sophomore",
                                        "Junior",
                                        "Senior",
                                        "Graduate Certificate Program")),
         CodedMajor = factor(CodedMajor, levels = c("Computer Science",
                                                    "Mathematical Sciences",
                                                    "Other Sciences",
                                                    "Other")))
```


## Now the Years Table Again

The class years are now in the customary order, as determined by the imposed factor structure. 

```{r}
students %>% 
  count(Year) %>% 
  mutate(percents = 100 * prop.table(n))
```


## And the Majors Table

Now the catch-all `Other` category is last, as is typically seen in summaries, because of the imposed structure. There are also *ordered factors* that have additional structure for higher level analyses. We do not need those for EDA.

```{r}
students %>% 
  count(CodedMajor) %>% 
  mutate(percents = 100 * prop.table(n))
```


## Recoding Factors

Sometimes we want to rename or combine some factor levels. The `fct_recode` function from the `forcats` package allows us to do this. The function `fct_collapse()` is another alternative. This time we will create a new variable `Year2`.

```{r}
library(forcats) # or library(tidyverse)
students <- students %>% 
  mutate(Year2 = fct_recode(Year,
                            FR = "Freshman",
                            SO = "Sophomore",
                            JR = "Junior",
                            SR = "Senior",
                            GR = "Graduate Certificate Program"))
str(students$Year2)
```


##

Assigning more than one level to the same new level name will combine the levels. Here all the majors besides `Mathematical Sciences` become one group.

```{r}
library(forcats) # or library(tidyverse)
students <- students %>% 
  mutate(CodedMajor2 = fct_recode(CodedMajor,
                                  Math = "Mathematical Sciences",
                                  Not_Math = "Computer Science",
                                  Not_Math = "Other Sciences",
                                  Not_Math = "Other"))
str(students$CodedMajor2)
```


## Use a Single `mutate()` Statement

In these examples, you could combine both the `fact_recode()` functions in the same `mutate()` statement you used to create your factors. That is another advantage of the `tidyverse` way. This example has  just `Year` due to limited slide space.

```{r}
students <- students %>% 
  mutate(Year = factor(Year, levels = c("Freshman",
                                        "Sophomore",
                                        "Junior",
                                        "Senior",
                                        "Graduate Certificate Program")),
         Year2 = fct_recode(Year,
                            FR = "Freshman",
                            SO = "Sophomore",
                            JR = "Junior",
                            SR = "Senior",
                            GR = "Graduate Certificate Program"))
```
