---
title: "EDA &#9654; Numerical Data<br>Center, Spread, Z-Scores"
author: "Author: Jill E. Thomley"
date: '`r format(Sys.time(), "%B %d, %Y @ %I:%M %p")`'
output: 
  ioslides_presentation:
    logo: images/logoASU.jpg
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(comment = "", message = FALSE, warning = FALSE)
library(tidyverse)
library(moments)
```


## Before We Begin...

These slides are not meant to be standalone information. You should take notes to flesh out the contents. I recommend that you create an R Markdown document where you can combine information and code from the slides and your own additional notes and explorations to make connections.

**Related Materials**

* Ch 2 of *Mathematical Statistics with Resampling and R, 2^nd^ Ed.*
* Ch 2 of [*Modern Dive*](https://moderndive.com/2-viz.html)
* Ch 3 of [*Modern Dive*](https://moderndive.com/3-wrangling.html)
* Ch 4 of [*Modern Dive*](https://moderndive.com/4-tidy.html)
* DataCamp [Introduction to the Tidyverse](https://www.datacamp.com/courses/introduction-to-the-tidyverse)


## How We Describe Distributions

**Categorical Variables**

* frequency / relative frequency

*Summaries based on counting how many times categories occur.*

**Quantitative Variables**

* shape
* center
* spread
* outliers

*Shape, center, and spread characterize the pattern of numerical observations. Outliers are deviations from the pattern.*


## Center

* mean
    * arithmetic mean ("average", _expected value_)
    * weighted mean
    * trimmed mean
    * geometric mean (n^th^ root of the product)
* median (50^th^ percentile/quantile)
* mode (we use this more with reference to shape)

Mean and median are both the "middle" of a given distribution, but the two summaries define middle or _center_ differently and may be very different values when calculated. If a distribution is perfectly symmetric, then mean = median.


## Spread (or *Variability*)

* variance (*variance* is one specific measure of *variability*)
    * population
    * sample
* standard deviation
* mean absolute deviation (mad)
* median absolute deviation (_also_ mad?)
* interquartile range (IQR)
* range

Summarizing the deviations of a set of values from some center point (like the mean) is a common technique in statistics. We will see this in many applications.


## Measures Based on Moments---Mean

In physics, a *moment of force* (or *torque*) is the turning effect of the force about a central point, as a function of distance from the point. Statistical moments are a similar idea.

mean (first moment)

* $X - \mu$ calculates the distance of point X from the mean $\mu$
* the mean $\mu$ is the value that makes $\Sigma(X - \mu)^1 = 0$ true for all points $X_i$ in a given distribution
* mean is the **center of mass** or *balance point* of a distribution

Think of the number line as a see-saw and the data points as objects sitting on the see-saw, exerting a downward force. The mean is the fulcrum that perfectly balances all the data points and keeps the see-saw level.


## Variance and Standard Deviation

variance (second moment)

$$M_2 = \sigma^2 = \frac{\Sigma(X - \mu)^2}{n}$$

The variance is the average *squared* distance of all points from the mean. Points that are farther away from $\mu$ will be weighted somewhat more heavily. Whatever units of measure $X$ has will be squared as well, so it is less useful for description.

$$\text{standard deviation} = \sqrt{\text{variance}} = \sigma $$

While not strictly mathematically true, it is helpful to think of SD as roughly the average distance of all points from the mean. Its units are also the same as the mean and the original data.


## Population vs. Sample Variance

The previous slide shows the formula for *population* variance. 

When computing sample variance, we make a small change to the formula so that it is an *unbiased* estimator of the population variance (i.e., in the long run sample variance will neither over-estimate nor under-estimate population variance).

$$\text{sample variance} = s^2 = \frac{\Sigma(x_i - \bar{x})^2}{n-1}$$ 

$$\text{sample standard deviation} = \sqrt{\text{sample variance}} = s$$

Recall we use different symbols for parameters versus statistics. If you are not yet familiar with the symbols for mean, variance, standard deviation, and sample size, this is a good time to learn!


## Z-Score (Standardized Score)

Mean and standard deviation have the same units as the data points they are summarizing (e.g., data in inches have a mean expressed in inches). 

However, standard deviation itself can be a unit of measure. A z-score tells us how many standard deviations above or below the mean a given point is located.

$$z = \frac{X - \mu}{\sigma} \text{ (population)}$$

$$z = \frac{x - \bar{x}}{s} \text{ (in a sample)}$$

The data units cancel out. A z-score of +2 would be interpreted as "two standard deviations above the mean".


## The Mathematics of Z-Scores

Why are z-scores *standardized* scores? The calculation performs a linear transformation (re-scaling) of the values, similar to when we convert degrees Fahrenheit to Celsius. Data points maintain their relative relationships to each other, we are just using a new number line scale as a reference point.

$(X - \mu)$ re-centers the data, and the new mean is always zero

dividing by $\sigma$ re-scales the number line, so the SD is always one

The act of dividing also cancels out original measurement units. We then talk about data points in terms of how many standard deviations they are above or below the mean. It helps compare variables that may originally have been on very different scales by using a common or *standard* scale.


## Why do We Need Z-Scores?

As noted, z-scores help us compare data and distributions that may be measures on different scales, the way we might convert measurements made in inches, feet, and meters to a common set of units to make more direct comparisons. 

Our traditional methods of estimation and inference also often rely on using standardized measures so that we can use these same methods in a broad variety of situations. In statistics, our data is an unknown, but the same way scientists calibrate their instruments, we make sure our statistical methods are reliable by validating and "calibrating" them mathematically.

In the future, one variation you see on z-scores will be *t*-scores, used for estimation and inference about means.


## Skewness and Kurtosis

skewness (based on the third moment, notice z-score similarity)

$$M_3 = \frac{\Sigma(X - \mu)^3}{n}$$

$$skewness = \frac{\Sigma(X - \mu)^3}{\sigma^3}$$

kurtosis (based on the fourth moment, notice z-score similarity)

$$M_4 = \frac{\Sigma(X - \mu)^4}{n}$$

$$kurtosis = \frac{\Sigma(X - \mu)^4}{\sigma^4}$$


## More on Skewness, Kurtosis, Moments

Skewness and kurtosis help to judge the *shape* of a distribution. 

* skewness is the degree of asymmetry in the distribution

* kurtosis is the flatness or peakedness of the distribution

We use these measures often when trying to judge whether or not we can consider a data to be approximately bell-shaped, since many of our traditional methods assume that data come from a population this characteristic shape.

We will talk about these measures in more depth when we get to plotting data, shape, and the normal distribution.

Outliers will affect all measures based on moments, so we need to check for them when we calculate these measures.


## Moment-Based Summaries in R

* `mean()`
* `var()`
* `sd()`
* `skewness()`
* `kurtosis()`

The `var()` and `sd()` functions calculate sample variance and standard deviation. How could you get population variance if your data were a population?

The `skewness()` and `kurtosis()` functions are found in the `moments` package, so you need to use `library(moments)`.

You should explore the syntax and output of these functions.


## Measures Based on Ordered Counts

* five-number-summary
    * minimum &mdash; smallest data point
    * first quartile (Q<sub>1</sub>) &mdash; 25% of data is below, 75% is above
    * median &mdash; 50% of data is below, 50% is above ("center")
    * third quartile (Q<sub>3</sub>) &mdash; 75% of data is below, 25% is above
    * maximum &mdash; largest data point
* interquartile range (IQR) &mdash; Q<sub>3</sub> - Q<sub>1</sub> (range of middle 50%)
* range &mdash; maximum - minimum (overall width of the dataset)
* percentiles / quantiles &mdash; the n*th* percentile or quantile has n% of the data below it (e.g., Q<sub>1</sub> is the 25th percentile of the dataset and the median is the 50th percentile)


## Count-Based Summaries in R

* `fivenum()`
* `min()`
* `median()`
* `max()`
* `IQR()`
* `range()`
* `diff(range())`
* `quantile()`

Again, you should explore the syntax, parameters, and output of each function to understand how it works.


## Some Data to Summarize

Read datasets [gasmileage.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv) and [happyface.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/happyface.csv) into RStudio. Remember you need `library(tidyverse)` for many of the functions we will use, including `read_csv()`. In coming slides they will be referred to by the following names.

* `mileage`: results of 100 EPA gas mileage tests for a particular model of car, in miles per gallon (mpg)

* `happyf`: results for an experiment where some servers drew happy faces on the back of  customer's restaurant checks and others did not, with the tip percentage they received

In `happyf` the sex of the server is recorded as a binary variable `male` or `female`, since it is an older dataset ([published in 1996](https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1559-1816.1996.tb01847.x)). Much of the data we encounter in real life is time-sensitive; we need to keep this in mind when we interpret/discuss it.


## Explore the Datasets

```{r, echo = FALSE}
mileage <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv")
happyf <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/happyface.csv")
```

```{r}
glimpse(mileage)
glimpse(happyf)
```


## Summarizing Using `dplyr`

You can add/remove summary functions as needed. The `na.rm` ignores any `NA` (missing) values, else you get `NA` as output.

```{r, eval = FALSE}
library(moments) # you only need this for skewness and kurtosis
datasetname %>% 
  summarize(n    = n(),
            xbar = mean(variablename, na.rm = TRUE),
            s    = sd(variablename, na.rm = TRUE),
            min  = fivenum(variablename)[1],
            Q1   = fivenum(variablename)[2],
            med  = fivenum(variablename)[3],                  
            Q3   = fivenum(variablename)[4],
            max  = fivenum(variablename)[5],
            iqr  = IQR(variablename, na.rm = TRUE),
            rng  = diff(range(variablename, na.rm = TRUE)),
            skew = skewness(variablename, na.rm = TRUE),
            kurt = kurtosis(variablename, na.rm = TRUE))
```


## About the `fivenum()` Function

Most summary functions like `mean()` and `sd()` produce only a single value as output. However, `fivenum()` produces a vector of length 5. The [ ] notation accesses the correct location in the output vector, since each step of a `summarize()` function must only output a single value.

```{r}
fivenum(mileage$mpg) # output in order is min, Q1, med, Q3, max
fivenum(mileage$mpg)[3] # the output is location 3, the median
```

Also, `range()` outputs (min, max); we have to take the `diff()`.


## Summarizing `mpg`

I removed `skewness()` and `kurtosis()` here, but added `sum()`. I also know for sure that there are no outliers in this dataset, so I removed the `na.rm = TRUE` statements for simplicity. 

```{r, eval = FALSE}
mileage %>% 
  summarize(n     = n(),
            xbar  = mean(mpg),
            s     = sd(mpg),
            min   = fivenum(mpg)[1],
            Q1    = fivenum(mpg)[2],
            med   = fivenum(mpg)[3],                  
            Q3    = fivenum(mpg)[4],
            max   = fivenum(mpg)[5],
            iqr   = IQR(mpg),
            rng   = diff(range(mpg)),
            total = sum(mpg))
```
 

## The Output is a `tibble`

Both the input and output are datasets, but the components of the output are the summaries computed from the input dataset. Each summary measure is a variable, with a single observation. You can `filter()` and `select()` and otherwise manipulate this `tibble`, especially if you save it as an object.

```{r echo = FALSE}
# I know these data have no NA values
mileage %>% 
  summarize(n     = n(),
            xbar  = mean(mpg),
            s     = sd(mpg),
            min   = fivenum(mpg)[1],
            Q1    = fivenum(mpg)[2],
            med   = fivenum(mpg)[3],                  
            Q3    = fivenum(mpg)[4],
            max   = fivenum(mpg)[5],
            iqr   = IQR(mpg),
            rng   = diff(range(mpg)),
            total = sum(mpg))
```

Where is the *center* of this distribution? How *spread* out is it? Keep in mind the units of measure when discussing the data and its summaries in context. 

We need to use visualizations to check for *shape* and *outliers*.


## 

```{r}
mpg_summaries <- mileage %>% 
  summarize(n     = n(),
            xbar  = mean(mpg),
            s     = sd(mpg),
            min   = fivenum(mpg)[1],
            Q1    = fivenum(mpg)[2],
            med   = fivenum(mpg)[3],                  
            Q3    = fivenum(mpg)[4],
            max   = fivenum(mpg)[5],
            iqr   = IQR(mpg),
            rng   = diff(range(mpg)),
            total = sum(mpg))
mpg_summaries %>% 
  select(n, xbar, s) # remember, select chooses columns (variables) 
```


## Summarizing by Groups Using `dplyr`

Use `group_by()` before the `summarize()` function to group the observations into subgroups by one or more other variables of interest. Here I use two, since the original experiment analyzed both together. 

```{r, eval = FALSE}
happyf %>% 
  group_by(Sex, Message) %>% 
  summarize(n    = n(),
            xbar = mean(TipPct),
            s    = sd(TipPct),
            min  = fivenum(TipPct)[1],
            Q1   = fivenum(TipPct)[2],
            med  = fivenum(TipPct)[3],                  
            Q3   = fivenum(TipPct)[4],
            max  = fivenum(TipPct)[5])
```
 

## Output for Grouped Summaries

The output is a tibble with four observations, or one for each group. We could use a `filter()` to choose only certain rows after summarizing.

```{r echo = FALSE}
happyf %>% 
  group_by(Sex, Message) %>% 
  summarize(n    = n(),
            xbar = mean(TipPct, na.rm = TRUE),
            s    = sd(TipPct, na.rm = TRUE),
            min  = fivenum(TipPct)[1],
            Q1   = fivenum(TipPct)[2],
            med  = fivenum(TipPct)[3],                  
            Q3   = fivenum(TipPct)[4],
            max  = fivenum(TipPct)[5])
```

You will usually see more decimal places in RStudio versus when you `knit()`. You can also format output using `kable()` styling. 


## 

```{r echo = FALSE}
happyfsummaries <- happyf %>% 
  group_by(Sex, Message) %>% 
  summarize(n    = n(),
            xbar = mean(TipPct, na.rm = TRUE),
            s    = sd(TipPct, na.rm = TRUE),
            min  = fivenum(TipPct)[1],
            Q1   = fivenum(TipPct)[2],
            med  = fivenum(TipPct)[3],                  
            Q3   = fivenum(TipPct)[4],
            max  = fivenum(TipPct)[5])
```

```{r}
# using summaries output that I saved as the object happyfsummaries
library(knitr)
kable(happyfsummaries, digits = 2) # digits = decimal places
```

<br>

Can you describe the center and spread for the four groups and make comparisons between them?


## The `quantile()` Function

Within the `quantile()` function, you can specify one or more quantiles to compute. Here we are finding the deciles of the `mpg` distribution. Creating vectors with functions like `seq()` can often make programming more efficient.

```{r}
quantile(mileage$mpg, probs = c(.1, .2, .3, .4, .5, .6, .7, .8, .9))

quantile(mileage$mpg, probs = seq(from = 0.1, to = 0.9, by = 0.1))
```


## Z-Scores with `dplyr`

Recall from earlier the formula for z-scores. Here we create a new variable that contains the z-score for every `mpg` value.

```{r}
mileage <- mileage %>% 
  mutate(zscores = (mpg - mean(mpg)) / sd(mpg))
head(mileage)
```


## Better Way with `scale()` Function

The `scale()` function is better here, because you can also use it with `group_by()` to get z-scores within groups.

```{r}
mileage <- mileage %>% 
  mutate(zscores = scale(mpg))
head(mileage)
```


## More on the `scale()` Function

The `scale()` function attaches some additional information to the `zscores` column. Note that the attributes `center` and `scale` are the mean and standard deviation of `mpg`.


```{r}
str(mileage)
```

