---
title: "Describing Distributions<br>&#9654; Quantitative Data"
author: "Author: Jill E. Thomley"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: ioslides_presentation
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(comment="", message=FALSE, warning=FALSE)
```

## Describing Distributions

categorial variables

* frequency
* relative frequency

__numerical variables__

* __shape__
* __center__
* __spread__
* __outliers__

Shape, center, and spread characterize the pattern of values or observations. Outliers are deviations from the pattern.


## Shape

* how many modes?
    * unimodal
    * bimodal
    * multimodal
* symmetric or skewed?
    * right skewed
    * left skewed
* skewness (magnitude of asymmetry)
* kurtosis (conformation of peak and tails)

We rely heavily on plots to assess shape. Summary measures like skewness and kurtosis should be interpeted _with_ plots.


## Plots

* dotplot
    * geom_dotplot()
* histogram
    * geom_histogram()
* density plot
    * geom_density()
* empirical cumulative distribution plot
    * stat_ecdf()
* boxplot
    * geom_boxplot()
    
    
## Plotting Some Data

Load [gasmileage.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv) into RStudio and let's make some plots! 

```{r, echo=FALSE}
mileage <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv")
```

```{r}
library(dplyr)
glimpse(mileage)
```


## Basic Dotplot

```{r, fig.height=4}
library(ggplot2)
ggplot(mileage, aes(x = mpg)) + geom_dotplot()
```


## Basic Histogram

```{r, fig.height=4.3}
ggplot(mileage, aes(x = mpg)) + geom_histogram()
```


## Basic Density

```{r, fig.height=4.3}
ggplot(mileage, aes(x = mpg)) + geom_density()
```


## Density Over Histogram

```{r, fig.height=3.7}
ggplot(mileage, aes(x = mpg)) + 
  geom_histogram(aes(y=..density..)) + 
  geom_density()
```


## Empirical Cumulative Distribution

```{r, fig.height=4.3}
ggplot(mileage, aes(x = mpg)) + stat_ecdf()
```


## Density and ECDF

```{r, fig.height=3.7}
ggplot(mileage, aes(x = mpg)) + 
  geom_density() + 
  stat_ecdf()
```


## Skewness and Kurtosis

```{r}
library(moments)
skewness(mileage$mpg)
kurtosis(mileage$mpg)
kurtosis(mileage$mpg) - 3
```


## Center

* mean
    * arithmetic mean ("average", _expected value_)
    * geometric mean
    * weighted mean
    * trimmed mean
* median
* mode (we use this more with reference to shape)

Mean and median are both the "middle" of a given distribution, but the two summaries define middle differently.


## Spread

* variance
    * population
    * sample
* standard deviation
* mean absolute deviation (mad)
* median absolute deviation (_also_ mad?)
* interquartile range (IQR)
* range

Summarizing the deviations of a set of values from some center point is a common technique in statistics. 


## Outliers

Coming soon!


