---
title: "Describing Distributions<br>&#9654; Quantitative Data"
author: "Author: Jill E. Thomley"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: ioslides_presentation
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(comment="", message=FALSE, warning=FALSE)
```

## Describing Distributions

categorial variables

* frequency
* relative frequency

__numerical variables__

* __shape__
* __center__
* __spread__
* __outliers__

Shape, center, and spread characterize the pattern of values or observations. Outliers are deviations from the pattern.


## Shape

* how many modes?
    * unimodal
    * bimodal
    * multimodal
* symmetric or skewed?
    * right skewed
    * left skewed
* skewness (quantification of asymmetry)
* kurtosis (conformation of peak and tails)

We rely heavily on plots to assess shape. Summary measures like skewness and kurtosis should be interpeted _with_ plots.


## Plots for Assessing Shape

* histogram
    * [geom_histogram()](https://ggplot2.tidyverse.org/reference/geom_histogram.html)
* density plot
    * [geom_density()](https://ggplot2.tidyverse.org/reference/geom_density.html)
* empirical cumulative distribution plot
    * [stat_ecdf()](https://ggplot2.tidyverse.org/reference/stat_ecdf.html)
* boxplot
    * [geom_boxplot()](https://ggplot2.tidyverse.org/reference/geom_boxplot.html)

See AsULearn Statistics Terms & Concepts glossary and course textbooks for more information on plots.

    
## Plotting Some Data

Load [gasmileage.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv) and [oldfaithful.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/oldfaithful.csv) into RStudio. Let's plot! 

```{r, echo=FALSE}
mileage <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv")
geyser  <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/oldfaithful.csv")
```

```{r}
library(dplyr)
glimpse(mileage)
glimpse(geyser)
```


## Basic Histogram

```{r, fig.height=4}
library(ggplot2)
ggplot(mileage, aes(x = mpg)) + geom_histogram()
```


## &nbsp;

```{r, fig.height=4.3}
ggplot(geyser, aes(x = wait_time)) + geom_histogram()
```


## Rules of Thumb for Number of Bins

* between 5 and 15 bins
* between 5 and 20 bins
* number of bins near the square root of the sample size n
* Sturge’s rule, number of bins around 1 + 3.322 × log(n)
* Freedman-Diaconis rule, about 2 × IQR × n^(−1/3) bins

Finding the right binwidth is an interative process. You can set an approximate number of bins, a specific bin width, or exact boundaries you want for the bins.

The default number of bins is 30.

https://ggplot2.tidyverse.org/reference/geom_histogram.html


## Basic Density

```{r, fig.height=4.3}
ggplot(mileage, aes(x = mpg)) + geom_density()
```


## &nbsp;

```{r, fig.height=4.3}
ggplot(geyser, aes(x = wait_time)) + geom_density()
```


## Density Over Histogram

```{r, fig.height=3.7}
ggplot(mileage, aes(x = mpg)) + 
  geom_histogram(aes(y=..density..)) + 
  geom_density()
```


## &nbsp;

```{r, fig.height=3.7}
ggplot(geyser, aes(x = wait_time)) + 
  geom_histogram(aes(y=..density..)) + 
  geom_density()
```


## Empirical Cumulative Distribution

```{r, fig.height=4.3}
ggplot(mileage, aes(x = mpg)) + stat_ecdf()
```


## Density and ECDF

```{r, fig.height=3.7}
ggplot(mileage, aes(x = mpg)) + 
  geom_density() + 
  stat_ecdf()
```


## Skewness and Kurtosis

```{r}
library(moments)
skewness(mileage$mpg)
kurtosis(mileage$mpg)
kurtosis(mileage$mpg) - 3
```


## More TBA

Additional slides will be added!