---
title: "Hypothesis Testing<br>&#9654; Means or Proportions"
author: "Author: Jill E. Thomley"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: ioslides_presentation
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(comment="", message=FALSE, warning=FALSE)
library(dplyr)
library(ggplot2)
```

## Before You Begin

Complete the following video lessons on AsULearn. They will introduce you to the logical framework of hypothesis testing, asymptotic tests for assessing the value of a single population mean or proportion using sample data, as well as comparing two population means or proportions.

* Theory and Concepts of Hypothesis Testing
* Classical Hypothesis Testing for One Mean
* Classical Hypothesis Testing for Two Means
* Classical Hypothesis Testing for Proportions

Also read the appropriate sections of _Chihara_ Chapter 3. Class time will be used to supplement these materials, introduce the concepts and methods of permutation tests, and help develop connections between the mathematics and programming.  


## Comments on Notation

Statistical notation can vary, since many disciplines have made contributions to statistical theory and practice. For means, we most often see the population mean represented by $\mu$ and the sample mean represented by $\bar{x}$. The textbook, the lesson videos, and these slides all follow this convention.

For proportions, one common notation is to use $\pi$ to represent the population proportion with and $p$ for the sample proportion (Greek versus Roman letters, like $\mu$ and $\bar{x}$). These slides use that convention. However, the lesson videos use another common system where the population proportion is $p$ and the sample proportion is $\widehat{p}$. The textbook seems to use a mix of notations.

In any context, make sure you know whether $p$ is being used as the population parameter or sample statistic.


# Comparing Two Means

## Null and Alternative Hypotheses

The hypotheses for a test comparing means are written in terms of the population means $\mu$.

null hypothesis

$$H_0: \mu_1 = \mu_2 \text{ or } H_0: \mu_1 - \mu_2 = 0$$

alternative hypothesis (two-tailed, lower-tailed, or upper-tailed)

$$H_a: \mu_1 \neq \mu_2 \text{ (or } H_a: \mu_1 - \mu_2 \neq 0)$$

$$H_a: \mu_1 < \mu_2 \text{ (or } H_a: \mu_1 - \mu_2 < 0)$$

$$H_a: \mu_1 > \mu_2 \text{ (or } H_a: \mu_1 - \mu_2 > 0)$$


## Traditional Test for Two Means

assumptions

* independent simple random samples from the populations
* normally distributed populations (or large enough samples)

test statistic (unpooled)

$$t=\frac{\bar{x}_1-\bar{x}_2}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}}$$

The p-value is calculated using a _t_ distribution with $\nu$ degrees of freedom, where $\nu$ depends on the sample sizes.

In R, we use `t.test()` to perform a test for two means.


## Comparing Two Means Example

Do brand name or generic batteries last longer?  ([batterylife.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/batterylife.csv))

```{r, echo=FALSE}
battery <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/batterylife.csv")
```

```{r}
str(battery)
summaries <- battery %>% 
  group_by(type) %>% 
  summarize(n = n(), xbar = mean(hours), stdv = sd(hours))
print.data.frame(summaries, row.names = FALSE)
```

## Are there any outliers in the data?

```{r, echo=FALSE}
ggplot(battery, aes(x = type, y = hours)) + geom_boxplot()
```


## Are the data normally distributed?

```{r, echo=FALSE}
ggplot(battery, aes(sample = hours, color = type)) + stat_qq() + stat_qq_line()
```


## The _t_ Distribution

As degrees of freedom $\nu \rightarrow \infty$, the $t$ distribution converges to the standard normal distribution.

```{r, echo=FALSE, fig.height=4}
ggplot(NULL, aes(x = c(-8, 8))) +
  stat_function(fun = dt, args = list(df = 2), color = "blue") +
  stat_function(fun = dt, args = list(df = 10), color = "darkgreen") +
  stat_function(fun = dnorm, args = list(0, 1), color = "red") +
  geom_segment(aes(x = 5, y = 0.20, xend = 6, yend = 0.20), color = "blue") +
  geom_segment(aes(x = 5, y = 0.18, xend = 6, yend = 0.18), color = "darkgreen") +
  geom_segment(aes(x = 5, y = 0.16, xend = 6, yend = 0.16), color = "red") +
  geom_text(aes(x = 6.2, y = 0.20, label = "2 DF"), hjust = "left", color = "blue") +
  geom_text(aes(x = 6.2, y = 0.18, label = "10 DF"), hjust = "left", color = "darkgreen") +
  geom_text(aes(x = 6.2, y = 0.16, label = "N(0,1)"), hjust = "left", color = "red") +
  scale_x_continuous(breaks = seq(-8, 8, 1)) +
  labs(x = "x", y = "f(x)")
```


## Tests for Two Means

Open the R Markdown file [Testing Two Means](https://raw.githubusercontent.com/STAT-JET-ASU/StatisticalDataAnalysis1/master/TestingTwoMeans.Rmd) to continue. We will assess the hypotheses three different ways.

* traditional asymptotic t-test
* complete permutation test
* permutation resampling test

Pay attention to the assumptions of each test, the test statistic used, and the way in which the p-value is calculated.


# Comparing Proportions

## Null and Alternative Hypotheses

The hypotheses for a test comparing proportions are written in terms of the population proportions $\pi$.

null hypothesis

$$H_0: \pi_1 = \pi_2 \text{ or } H_0: \pi_1 - \pi_2 = 0$$

alternative hypothesis (two-tailed, lower-tailed, or upper-tailed)

$$H_a: \pi_1 \neq \pi_2 \text{ (or } H_a: \pi_1 - \pi_2 \neq 0)$$

$$H_a: \pi_1 < \pi_2 \text{ (or } H_a: \pi_1 - \pi_2 < 0)$$

$$H_a: \pi_1 > \pi_2 \text{ (or } H_a: \pi_1 - \pi_2 > 0)$$


## Traditional Two-Proportion Test

assumptions

* independent simple random samples from the populations
* large enough sample sizes for the normal approximation to the binomial to be valid (one rule is np &ge; 15 and n(1-p) &ge; 15)

test statistic 

$$z=\frac{p_1-p_2}{\sqrt{p(1-p)(\frac{1}{n_1}+\frac{1}{n_2})}}$$

The p-value is calculated using the standard normal distribution.

In R, we use `prop.test()` to conduct a test for two proportions.


## Comparing Proportions Example

Does a magnetic pulse affect navigation?  ([homingpigeons.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/homingpigeons.csv))

```{r, echo=FALSE}
homing <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/homingpigeons.csv")
```

```{r}
str(homing)
summaries <- homing %>% 
  group_by(group) %>% summarize(n = n(), 
                                p_hat_y = mean(home == "no"), 
                                p_hat_n = mean(home == "yes"))
```

```{r, echo=FALSE}
print.data.frame(summaries, row.names = FALSE)
```

## Visualize the Proportions

```{r, echo=FALSE}
ggplot(homing, aes(x = group, fill = home)) + geom_bar(position = "fill") + ylab("proportion")
```


## Are np &ge; 15 and n(1&minus;p) &ge; 15?

Recall the summaries we computed. We can use them to check the assumption by multiplying the appropriate vectors.

```{r, echo=FALSE}
print.data.frame(summaries, row.names = FALSE)
```

```{r}
summaries$n * summaries$p_hat_y >= 15
summaries$n * summaries$p_hat_n >= 15
```


## Tests for Two Proportions

Continue using R Markdown file [Testing Two Proportions](https://raw.githubusercontent.com/STAT-JET-ASU/StatisticalDataAnalysis1/master/TestingTwoProportions.Rmd). We will assess the hypotheses three different ways.

* traditional test as implemented in R
    + the relationship between Z and $\chi^2$ distributions
    + continuity correction for normal approximation
* permutation resampling test

Pay attention to the assumptions of each test, the test statistic used, and the way in which the p-value is calculated.


# Other Parameters

## Two Medians and Two Variances

* Let $\theta_1$ and $\theta_2$ be the medians of two populations, estimated by sample medians $\widehat{\theta}_1$ and $\widehat{\theta}_2$. Test whether the population medians are equal or not using permutation resampling.

* Let $\sigma^2_1$ and $\sigma^2_2$ be the variances of two populations, estimated by sample variances $S^2_1$ and $S^2_2$. Test whether the population variances are equal or not using permutation resampling.<br><br>The standard test statistic for two variances is the ratio $\frac{S^2_1}{S^2_2}$.<br><br>Compare resampling results to the results of `var.test()`.











