---
title: "An Introduction to the Bernoulli Distribution"
subtitle: "Companion `R` Code for JB Statistics Video Lesson"
author: "Created by Jill E. Thomley"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4
    self_contained: false
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

***
```{r, echo = FALSE}
vembedr::embed_youtube(
id = "bT1p5tJwn_0",
width = NULL,
height = 400,
ratio = c("16by9", "4by3"),
frameborder = 0,
allowfullscreen = TRUE,
query = NULL
)
```
***

### **Packages Used**

<details>
```{r loadpackages}
library(tidyverse)
library(purrr)
```
</details>

### **The Bernoulli Distribution**

A Bernoulli trial is a single random event that has only two possible outcomes. These are often called "success" and "failure". Classic examples include one coin flip where a head is considered a success and a tail is considered a failure, or one roll of a standard 6-sided die where 1 is considered a success and any other number 2-6 is considered a failure.

$$X \sim Bernoulli(p)$$

    x = number of successes = {0, 1}
    p = probability of success
    1 - p = probability of failure

$$f(x) = P(X = x) = p^x(1-p)^{n-x}$$

$$F(x)=P(X \leq x) = \sum\limits_{i=0}^x f(x) = \sum\limits_{i=0}^x p^i(1-p)^{n-i}$$

$$E[X] = p \text{ and } Var[X] = p(1-p)$$

### **Simulating Bernoulli Trials**

You can simulate Bernoulli trials using the function `rbernoulli(n, p)` from the purrr package. The parameter $n$ in this case is the number of times you want to repeat the Bernoulli process. For example, suppose we want twenty flips of a fair coin. The output is a logical vector with `TRUE` for a success and `FALSE` for a failure.

```{r}
(fifty_flips <- rbernoulli(50, 0.5))
```

The mean and variance of the sample will approximate the theoretical values of E[X] = 0.5 and Var[X] = 0.25. As a reminder, we can perform mathematical operations on logical vectors because `TRUE` is stored as `1` and `FALSE` as `0`.

```{r}
mean(fifty_flips)
var(fifty_flips)
```


Another (and better) way to perform this kind of simulation is with the `rbinom()` function. You will learn more about the Binomial distribution and its relationship to the Bernoulli in another video. 

***
```{r}
sessionInfo()
```
