---
title: "Binomial/Not Binomial Examples"
subtitle: "Companion `R` Code for JB Statistics Video Lesson"
author: "Created by Jill E. Thomley"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output:
  html_document:
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4
    self_contained: false
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

***
```{r, echo = FALSE}
vembedr::embed_youtube(
id = "UJFIZY0xx_s",
width = NULL,
height = 400,
ratio = c("16by9", "4by3"),
frameborder = 0,
allowfullscreen = TRUE,
query = NULL
)
```
***

### **Packages Used**

<details>
```{r loadpackages}
# no packages are required
```
</details>

### **The Binomial PMF**

The binomial distribution is a discrete probability distribution. It mathematically models the *number of successes* in a sequence of independent Bernoulli trials. The probability for a given number of successes $x$ is found using the *probability mass function* (pmf).

$$X \sim Bin(n,p)$$

    x = number of successes = {0, 1, 2, ..., n}
    n = number of trials
    p = probability of success on a single trial

$$f(x) = P(X = x) = {n \choose x}p^x(1-p)^{n-x}$$

### **Binomial Examples**

#### Chance of Fetal Down Syndrome

The probability a randomly selected 40-year-old pregnant woman is carrying a fetus with Down syndrome is approximately 0.01. Suppose we have an independent random sample of 25 40-year-old women. What is the probability exactly 2 women are carrying a fetus with Down syndrome?

* Trials are independent
* Fixed number of trials = 25
* Constant chance of success = 0.01
* We are counting the number of successes


$$X \sim Bin(25,0.01)$$

Solve for $x = 2$.

$$f(2) = P(X = 2) = {25 \choose 2}(0.01)^2(1-0.01)^{25-2} = \frac{25!}{2!\times 23!}(0.01)^2(0.99)^{23}$$

```{r}
dbinom(2, 25, 0.01)
```

#### Chance of 3 or 4 on a Die Roll

Suppose an ordinary six-sided die is rolled 16 times. What is the probability that on exactly 5 of the rolls, either a 3 or 4 comes up?

* Trials are independent
* Fixed number of trials = 16
* Constant chance of success = 2/6
* We are counting the number of successes

$$X \sim Bin(16, 2/6)$$

Solve for $x = 5$.

$$f(5) = P(X = 5) = {16 \choose 5}\left(\frac{2}{6}\right)^5\left(1-\frac{2}{6}\right)^{16-5}=\frac{16!}{5!\times 11!}\left(\frac{2}{6}\right)^5\left(\frac{4}{6}\right)^{11}$$
```{r}
dbinom(5, 16, 1/3)
```

***
```{r}
sessionInfo()
```
