---
title: "Project 2: Estimation of Mercury in Fish"
author: "PUT YOUR NAME HERE"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r globalopts, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

### **Packages/Data**

Load the packages and dataset here. Use `glimpse` to visualize the structure of the dataset.

Packages Used

```{r}

```

Fish Mercury Data

```{r}

```

***
### **Problem**

Nearly all fish contain some trace of mercury. For most people, this is not a serious health concern, but higher levels of mercury can harm the developing nervous systems of fetuses or young children and its effects when consumed can be cumulative. The Food and Drug Administration (FDA) and the Environmental Protection Agency (EPA) advise women who may become pregnant, pregnant women, nursing mothers, and young children to avoid some types of fish and eat others that are known to be lower in mercury. The EPA has [guidelines](https://www.epa.gov/fish-tech/epa-fda-fish-advice-technical-information) about how much is considered safe to eat. In general, the best choice is considered to be fish that have 15 µg/g. 

Commercial fish are regularly tested for mercury, but wild-caught fish are more difficult to regulate. Officials periodically sample local sources to assess the current level of risk from fish caught there. The dataset [FishMercury](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/FishMercury.csv) contains data for a sample of one frequently-consumed species of fish caught in a Minnesota lake. Mercury measurements are in parts per million (1 ppm = 1 µg/g). The data are also available in the `resampledata` package.

#### **Part A**

Compute the sample size, mean, standard deviation, median, and 5% trimmed mean for the sample. Also create a boxplot.

```{r}

```

```{r}

```

#### **Part B**

Compute a bootstrap 95% percentile interval to estimate the true mean level of mercury for this species of fish in the sampled lake.

```{r}

```

#### **Part C**

A common rule for determining when we can use the central limit theorem for the sampling distribution of the mean is n ≥ 30. Since that is true in this instance, compute a traditional 95% confidence interval to estimate the true mean level of mercury for this species of fish in the sampled lake. 

```{r}

```

#### **Part D**

Remove any outliers in the data and re-compute a bootstrap 95% percentile interval to estimate the true mean level of mercury for this species of fish in the sampled lake.

```{r}

```

#### **Questions**

Which of your three intervals is the most valid and appropriate to use, given the sample and the assumptions of each method? Explain.

**ANSWER:** 

Based on the interval you consider to be the most valid, can we confidently say that the true average level of mercury is less than 15 µg/g? Explain.

**ANSWER:** 


***
### Session Info

**Names of Collaborators**:

```{r}
sessionInfo()
```

