---
title: "Project 2: Binomial Problem"
author: "PUT YOUR NAME HERE"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r globalopts, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

### **Packages**

Load the packages here.

```{r}

```

***
### **Problem**

A professor gives her class a multiple-choice quiz with ten questions. Each question has four possible answers. A student who did not study for the quiz randomly chooses an answer for each question. Let X be the number of questions he or she gets correct on the quiz. This process can be modeled with a binomial distribution. 

#### **Part A**

Compute the expected value and variance of X.

```{r}

```

#### **Part B**

Create a plot of the probability mass function (pmf) of X.

```{r}

```

#### **Part C**

What is the probability that the student passes the quiz, if 60% correct is passing?

```{r}

```

#### **Part D**

Suppose that, on each question, the student can eliminate two of the possible answers as being definitely wrong. After eliminating them, he or she then randomly chooses between the two answers that are left. What is the probability that the student passes the quiz, if 60% correct is passing?

```{r}

``` 

#### **Questions**

One strategy people often say to use on multiple choice tests is to always choose "C" if you have to guess. Suppose the student in this problem used that method. Could we still model the probability with a binomial distribution? Is there anything else you would need to know about the quiz to decide? Discuss. 

**ANSWER:** 

Suppose the student could eliminate two of the possible answers on some questions but not on others. Could we still model the probability with a binomial distribution? Discuss.

**ANSWER:** 


***
### Session Info

**Names of Collaborators**:

```{r}
sessionInfo()
```

