---
title: "Project 2: Smallpox Hypothesis Test"
author: "PUT YOUR NAME HERE"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r globalopts, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

### **Packages/Data**

Load the packages and dataset here. Use `glimpse` to visualize the structure of the dataset.

Packages Used

```{r}

```

Smallpox Data

```{r}

```

***
### **Problem**

In the late 1700s, smallpox killed about one in four children born in London. Scientists knew that inoculations (i.e., vaccines) were effective in reducing mortality, but no one was sure how they worked. Doctors often prescribed tonics to prepare patients before vaccination. Dr. William Watson became the physician for a large London orphanage 1762. Administrators were concerned about rapid spread of disease in crowded conditions and ordered vaccinations for all of the children who lived there. Watson used this chance to investigate the effectiveness of vaccine pre-treatments. This is considered to be one of the first historical examples of the kind of research we now call a clinical trial. 

In his first experiment, Watson divided thirty-one children into three groups. 

* pre-vaccination treatment of mercury and jalap (a strong laxative/purgative)
* pre-vaccination treatment of senna and syrup of roses (a mild laxative)
* no pre-vaccination treatment (control group)

Watson observed and counted the number of smallpox lesions that appeared on each child. In general, more pocks is a sign of a more severe infection and greater chance of mortality. The data from Watson's experiment can be found in the file [smallpox.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/smallpox.csv). (Note: It was perfectly acceptable to conduct experiments like this throughout much of human history, but in the present day this would be considered unethical, especially with children as subjects.)

#### **Part A**

Reduce the dataset to extract data for Watson's first experiment only. Create a new variable that classifies the pretreatment group as either "control" or "laxative" versus the three original groups.

```{r}

```

#### **Part B**

Compute n, mean, and standard deviation for number of pocks that appeared in the `control` and `laxative` groups. Also create a boxplot to compare the distributions of the two groups.

```{r}

```

#### **Part C**

Conduct a permutation resampling test to determine whether there was a difference in the number of pocks between the control and laxative groups, on average. Use $\alpha$ = 0.05. Be sure to specify your hypotheses.

Ho:

Ha:

```{r}

```

Do you reject Ho? 

ANSWER: 

#### **Questions**

Based on your decision about Ho in Part C, which type of error could you have made, Type I or Type II? Explain.

**ANSWER:** 

Discuss the results of Watson's experiment. Did using laxatives as a vaccine pre-treatment impact the effectiveness of the vaccine, as measured by the number of pocks? Discuss using your exploratory data analysis and hypothesis test. 

**ANSWER:** 


***
### Session Info

**Names of Collaborators**:

```{r}
sessionInfo()
```

