---
title: "Hypothesis Testing Problem Session"
author: "Jill Thomley"
date: "04/04/2018"
output: html_document
---

Note: before conducting any kind of hypothesis test, it is a good idea to perform exploratory data analyses (as in Chapter 2) to become familiar with the distributions of your variables and any anomalies the data may contain. 

**Problem 1**

Consider the Flight Delays Case Study in Chapter 1 of the textbook. For each test, specify your null and alternative hypotheses, and if you reject Ho, discuss the results in the context of the problem.

```{r}
FD <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datafiles/master/Chihara/FlightDelays.csv")
```

A) The data contains flight delays for two airlines, American Airlines (AA) and United Airlines (UA). Are the mean delay times different for the two airlines? Conduct both a permutation test and a traditional test. Are the results consistent? What assumptions must be met in order for each test to be valid?

```{r}

```

B) Consider the proportion of times that each carrier's flight were delayed more than 20 minutes. Is the difference in proportions statistically significant? Conduct a permutation test.

```{r}

```

C) Is the variance in delay times for United Airlines greater than the variance in delay times for American Airlines? Conduct a permutation test. Use variance_AA / variance_UA as your test statistic.

```{r}

```

D) Is being delayed more than 30 minutes dependent on destination? Conduct both a traditional hypothesis test and a permutation test.

```{r}

```


**Problem 2**

Consider the GSS 2002 case study in Chapter 1 of the textbook. Conduct both a traditional hypothesis and a permutation test for each question below. In each case, specify your null and alternative hypotheses, and if you reject Ho, discuss the results in the context of the problem.

```{r}
GSS <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datafiles/master/Chihara/GSS2002.csv")
```

A) Are support for gun laws (GunLaw) and views on government spending on the military (SpendMilitary) independent?

B) Are support for gun laws and political party (PolParty) independent?

B) Are support for gun laws and gender independent?

