---
title: "WED 03/14/2018 Assignment"
author: "YOUR NAME HERE"
date: "Updated `r Sys.Date()`"
output: html_document
---

```{r, echo=FALSE, message=FALSE}
# Additional packages
library(car)
library(dplyr)
library(ggplot2)
library(moments)

TESTDATA <- read.csv(file=url("https://raw.githubusercontent.com/STAT-JET-ASU/Datafiles/master/Instructor/TESTDATA.csv"))
```

<hr>

#### THE PROBLEM

The TESTDATA dataset contains eight variables. Two of the variables are samples from a Normal or approximately Normal distribution. One is a sample from a Uniform distribution. One is a sample from an Exponential distribution with a constant subtracted from each value. The other four are distributions we have not worked with. Complete the descriptive analysis described below to help you identify which distribution is which. Go to DISTRIBUTION IDENTIFICATION on AsULearn to input your identification for each of the eight distributions.

#### YOUR ANALYSES

Compute the mean, standard deviation, skewness, and kurtosis for each variable using `dplyr` piping and functions. The output should be a single tibble containing all four statistics. Then create five plots for each variable. Put each set of analyses in the appropriate section below.

* a histogram with 10 bins and a boxplot using `ggplot2` 
* a density plot and an ECDF plot using base code or `ggplot2` 
* a normal quantile-quantile plot using the `qqPlot` function from the `car` package 

The qqPlot function is similar to using `qqnorm` with `qqline`, but it also includes guidelines to assist in determining whether the points are close enough to the straight line representing theoretically normal data. 

##### Variable 1

```{r}
# Erase the comment tag in front of TESTDATA and complete the code

# TESTDATA %>% 
```

##### Variable 2

```{r}
# Erase the comment tag in front of TESTDATA and complete the code

# TESTDATA %>% 
```

##### Variable 3

```{r}
# Erase the comment tag in front of TESTDATA and complete the code

# TESTDATA %>% 
```

##### Variable 4

```{r}
# Erase the comment tag in front of TESTDATA and complete the code

# TESTDATA %>% 
```

##### Variable 5

```{r}
# Erase the comment tag in front of TESTDATA and complete the code

# TESTDATA %>% 
```

##### Variable 6

```{r}
# Erase the comment tag in front of TESTDATA and complete the code

# TESTDATA %>% 
```

##### Variable 7

```{r}
# Erase the comment tag in front of TESTDATA and complete the code

# TESTDATA %>% 
```

##### Variable 8

```{r}
# Erase the comment tag in front of TESTDATA and complete the code

# TESTDATA %>% 
```

<hr>

DONE!!!