---
title: "Digital Communication, Part 2"
author: "PUT THE NAMES OF ALL GROUP MEMBERS HERE"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    highlight: textmate
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "")
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r, echo=FALSE}
# load all packages and datasets here

```

<hr>

1) In Part 1, you analyzed error rates for a single-channel system and a TRC system. One way to reduce transmission errors in either system would be to improve the initial 1% error rate. How much would reducing the error rate improve the performance of the TRC system? Create a plot that shows error rates from 0% to 1% (in increments of 0.1) on the x-axis versus the corresponding probability that any single TRC block (e.g., 000) is interpreted correctly (y-axis). To what value do we have to reduce the initial error rate in order for the probability of a block being correctly transmitted to be at least 99.99%?

```{r}

```

ANSWER: 


2) A 1000 Mbit/s system transmits 10^9^ bits per second. Each bit has a 0.03% chance of being incorrectly received when we use TRC and all bits are independent. There are are 86,400 seconds in a day. If X is the number of incorrect bits in a single second, use `rbinom` to simulate the distribution of X for a single day. Assess your simulation results for normality using a density plot, ECDF plot, QQ plot, and goodness of fit test. What do you conclude?

Simulation:
```{r}

```

Ho: 

Ha: 

Goodness of Fit Test:
```{r}
# test code here
```

Conclusions:


<hr>

END PART 2