---
title: "Project 03"
author: "YOUR NAME HERE"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    highlight: textmate
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "")
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r, message=FALSE}
# load all packages and datasets here

```

<hr>

### Your Analyses

1) Compute the sample size, mean, standard deviation, median, and 5% trimmed mean for the sample using `summarize` in `dplyr`. Create a boxplot using `ggplot2`. You do not need a title or different axis labels.

```{r}

```


2) A common rule for determining when we can use the central limit theorem is n ≥ 30. Since that is true in this instance, compute a traditional 95% confidence interval to estimate the true mean level of mercury for this species of fish in the sampled lake. 

```{r}
 
```


3) Compute a bootstrap 95% percentile interval to estimate the true mean level of mercury for this species of fish in the sampled lake. Do not include any plots in this section.

```{r}

```


4) Assess your bootstrap sampling distribution of $\bar{x}$ for normality using a density plot, an ECDF plot, and a quantile-quantile plot. Include red reference curves for the theoretical normal distribution.

```{r}

```

5) Create an ordered factor variable called `safety` that categorizes fish in the sample as "Best", "Good", or "Avoid". Then create a barplot to display that variable. Include an informative title and axis labels.

```{r}

```


6) Compute a bootstrap 95% percentile interval to estimate the true proportion of this species of fish that fall into the "best" category. Do not include any plots in this section.

```{r}

```


7) Why are your intervals in (2) and (3) different? Is it appropriate to the CLT for these data? Expain using the results of (1) and (4). 

ANSWER:


8) Assuming this is a representative sample, can we confidently say a majority of all fish of this species in the lake have fewer than 0.15 µg/g of mercury? Why or why not?

ANSWER: 

<hr>

END PROJECT 3