---
title: "Vietnam Draft, Part 2"
author: "PUT THE NAMES OF ALL GROUP MEMBERS HERE"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    highlight: textmate
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "")
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r, echo=FALSE}
# load all packages and datasets here

```

<hr>

For all tests, write the null and alternative hypotheses, report the test statistic and the p-value, state whether or not you reject the null hypothesis, and then give your conclusion in terms of the original question.

1) Perform both a traditional and a permutation hypothesis to determine whether the mean draft number for the second half of the year and the mean draft number for the first half of the year are significantly different. If so, which mean is smaller?

Ho: 

Ha: 

Traditional Test:
```{r}
# test code here
```

Permutation Test:
```{r}
# test code here
```

Conclusions:


2) As noted above, the largest draft number called up for service in 1970 was 195. Perform a both a traditional and permutation hypothesis to determine whether the the second half of the year had a significantly larger proportion of draft numbers that were less than or equal to 195 compared to the first half of the year. 

Ho: 

Ha: 

Traditional Test:
```{r}
# test code here
```

Permutation Test:
```{r}
# test code here
```

Conclusions:


3) The overall distribution of draft numbers is discrete uniform. If the process was truly random, then the distribution of draft numbers in each half of the year should also be essentially uniform as well. Perform a goodness of fit test to determine whether draft numbers in the second half of the year are uniformly distributed.

Ho: 

Ha: 

Traditional Test:
```{r}
# test code here
```

Conclusions:


4) How do your analyses support the assertion that the draft was unfair? Why are they not _proof_ of unfairness?

ANSWER:


<hr>

END PART 2





