---
title: "Battery Life Example"
subtitle: "Permutation Resampling Hypothesis Test"
author: "Jill E. Thomley"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = "")
```

```{r}
library(tidyverse)
```

```{r}
battery <- read.csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/batterylife.csv"))
battery
```


### Descriptive Statistics

```{r}
stats <- battery %>% 
  filter(!is.na(hours)) %>% 
  group_by(type) %>% 
  summarize(xbar = mean(hours),
            stdv = sd(hours),
            n    = n())

print.data.frame(stats)

ggplot(battery, aes(x = type, y = hours)) +
  geom_boxplot() +
  coord_flip()

ggplot(battery, aes(sample = hours, color = type)) +
  geom_qq() +
  geom_qq_line()
```

<hr>

### Possible Hypotheses

#### Question 1: Do generic batteries last a longer amount of time than brand name batteries?

$$H_0: \mu_G = \mu_B \rightarrow \mu_G - \mu_B = 0$$

$$H_a: \mu_G > \mu_B \rightarrow \mu_G - \mu_B > 0$$


#### Question 2: Do generic batteries last a shorter amount of time than brand name batteries?

$$H_0: \mu_G = \mu_B \rightarrow \mu_G - \mu_B = 0$$

$$H_a: \mu_G < \mu_B \rightarrow \mu_G - \mu_B < 0$$


#### Question 3: Do generic batteries last a different amount of time than brand name batteries?

$$H_0: \mu_G = \mu_B \rightarrow \mu_G - \mu_B = 0$$

$$H_a: \mu_G \neq \mu_B \rightarrow \mu_G - \mu_B \neq 0$$

#### Test Statistic

The same test statistic is used for all three tests, since they all have the same null hypothesis.

$$\text{test statistic } = \bar{x}_G - \bar{x}_B$$

```{r}
teststat <- stats$xbar[2] - stats$xbar[1]
```

The observed value of the test statistic $\bar{x}_G - \bar{x}_B$ for these data is `r round(teststat, 2)`.

<hr>

### Permutation Testing

How many unique regroupings would there be for an exact permutation test?

```{r}
sprintf("The number of unique regroupings is %d.", choose(12, 6))
```


We will use permutation resampling rather than an exact permutation test.

```{r}
hours <- battery %>% filter(!is.na(hours)) %>% pull(hours)

total_n <- length(hours)
group_n <- stats$n[2]  

alpha <- .05

sims <- 10^5 - 1

randstat <- numeric(sims)

for (i in 1:sims) {
  index <- sample(total_n, group_n)
  randstat[i] <- mean(hours[index]) - mean(hours[-index])
}

ggplot(NULL, aes(x = randstat)) +
  geom_histogram(color = "gray", fill = "lightblue") +
  geom_vline(xintercept = teststat, color = "blue", linetype = "dashed") +
  labs(title = "Permutation Resampling Distribution for Test Statistic",
       subtitle = "vertical dashed line indicates the observed value of the test statistic") +
  theme_classic()
```

<hr>

### Computing P-Values

#### Question 1: Upper-Tail P-Value

$$H_a: \mu_G > \mu_B \rightarrow \mu_G - \mu_B > 0$$

```{r}
pval_upper <- (sum(randstat >= teststat) + 1) / (sims + 1)
decision <- ifelse(pval_upper <= alpha, "reject the null hypothesis", "not reject the null hypothesis")
pval_upper
decision
```

The p-value for this test is `r pval_upper`. Based on $\alpha$ = `r alpha`, we should `r decision`.


#### Question 2: Lower-Tail P-Value

$$H_a: \mu_G < \mu_B \rightarrow \mu_G - \mu_B < 0$$

```{r}
pval_lower <- (sum(randstat <= teststat) + 1) / (sims + 1)
decision <- ifelse(pval_lower <= alpha, "reject the null hypothesis", "not reject the null hypothesis")
pval_lower
decision
```

The p-value for this test is `r pval_lower`. Based on $\alpha$ = `r alpha`, we should `r decision`.


#### Question 3: Two-Tailed P-Value

$$H_a: \mu_G \neq \mu_B \rightarrow \mu_G - \mu_B \neq 0$$

```{r}
pval_upper <- (sum(randstat >= teststat) + 1) / (sims + 1)
pval_lower <- (sum(randstat <= teststat) + 1) / (sims + 1)
pval_2side <- 2 * min(pval_lower, pval_upper)
decision <- ifelse(pval_2side <= alpha, "reject the null hypothesis", "not reject the null hypothesis")
pval_2side
decision
```

The p-value for this test is `r pval_2side`. Based on $\alpha$ = `r alpha`, we should `r decision`.


