---
title: "Battery Life Example"
subtitle: "Permutation Resampling Hypothesis Test"
author: "Jill E. Thomley"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = "")
```

```{r}
library(tidyverse)
```

```{r}
draft <- read.csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/vietnamdraft.csv")) %>% 
  filter(draftyear == 1972) %>% 
  arrange(draftnumber)
draft
```


### Descriptive Statistics

```{r}
stats <- draft %>% 
  filter(!is.na(draftnumber)) %>% 
  group_by(halfyear) %>% 
  summarize(prop = mean(draftnumber <= 95),
            n    = n())

print.data.frame(stats)
```

<hr>

### Possible Hypotheses

#### Question 1: Is the proportion of numbers <= 195 greater in the second half of the year than the first?

$$H_0: \pi_{first} = \pi_{second} \rightarrow \pi_{first} - \pi_{second} = 0$$

$$H_a: \pi_{first} = \pi_{second} \rightarrow \pi_{first} - \pi_{second} \neq 0$$

```{r}
teststat <- stats$prop[1] - stats$prop[2]
teststat
```

<hr>

### Permutation Testing

How many unique regroupings would there be for an exact permutation test?

```{r}
choose(366, 182) # how many regroupings
factorial(366)   # how many orderings
```


We will use permutation resampling rather than an exact permutation test.

```{r}
draftnumber <- draft %>% filter(!is.na(draftnumber)) %>% pull(draftnumber)

total_n <- length(draftnumber)
group_n <- stats$n[1]  

alpha <- .05

sims <- 10^6 - 1

randstat <- numeric(sims)

for (i in 1:sims) {
  index <- sample(total_n, group_n)
  randstat[i] <- mean(draftnumber[index] <= 95) - mean(draftnumber[-index] <= 95)
}

ggplot(NULL, aes(x = randstat)) +
  geom_histogram(color = "gray", fill = "lightblue") +
  geom_vline(xintercept = teststat, color = "blue", linetype = "dashed") +
  labs(title = "Permutation Resampling Distribution for Test Statistic",
       subtitle = "vertical dashed line indicates the observed value of the test statistic") +
  theme_linedraw()
```

<hr>

### Computing P-Values

#### Lower-Tail P-Value

$$H_a: \mu_G < \mu_B \rightarrow \mu_G - \mu_B < 0$$

```{r}
pval_upper <- (sum(randstat >= teststat) + 1) / (sims + 1)
pval_lower <- (sum(randstat <= teststat) + 1) / (sims + 1)
pval_2side <- 2 * min(pval_lower, pval_upper)
decision <- ifelse(pval_2side <= alpha, "reject the null hypothesis", "not reject the null hypothesis")
pval_2side
decision
```

The p-value for this test is `r pval_lower`. Based on $\alpha$ = `r alpha`, we should `r decision`.