---
title: "Exam 2"
author: "YOUR NAME HERE"
date: "`r Sys.time()`"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
# load packages and datasets here

```

**NOTE: All plots should have appropriate titles and axis labels. If you need to use a calculated numerical answer in a later analysis, you should store it and call it back when needed, not type the value by hand. If you store a calculated answer, be sure to print it as well so I can see it. Answer the written questions where it says ANSWER.**

### PROBLEM 1

A February 2014 [Vox article](https://www.vox.com/2014/11/7/7169501/same-sex-marriage-marijuana-legalization-midterm-election-2014) addressed the intersection of U.S. states that (at the time) allowed gay marriage and some form of legal marijuana use. Were those two issues (yes or no on gay marriage, yes or no on marijuana use) independent of one another? (data: mmFEB2015.csv)

(A) Create a contingency table and a barplot to explore the relationship between these two variables. 

```{r}

```

(B) Conduct a traditional chi-square test of independence (without continuity correction) and a simulation-based independence test using R's chisq.test function, as well as a permutation test of independence.

```{r}

```

(C) Based on the permutation test, should you reject the null hypothesis at the &alpha; = 0.05 level of significance? If you reject the null hypothesis, discuss the relationship between support for gay marriage and support for marijuana. Include relevant statistics to support your discussion.

ANSWER:

(D) Are the results in (B) consistent with one another? What assumption must be met for the traditional test that is not required for the randomization or Monte Carlo tests?

ANSWER:


### PROBLEM 2

In 2002, the New York Yankees had the highest average team salary among all MLB teams. However, was it statistically larger than the average salary of the rest of the teams in the league combined? (data: mlb2002.csv)

(A) Use the `mutate` function in `dplyr` and the `ifelse` function to add a new variable to the dataset that identifies players according to whether they played for the New York Yankees (NYY) or some other team (OTH). Then use the `summarize` function in `dplyr` to find n, mean, and standard deviation for salary in each group. Also create a boxplot using `ggplot2` to compare the two groups. 

```{r}

```

(B) Conduct a permutation test to determine whether the Yankee's mean was statistically larger than the mean of the rest of the league. Be sure to state your conclusions.

```{r}

```

ANSWER:

(C) Use bootstrap resampling with 10,000 resamples to find a 95% bootstrap percentile confidence interval for the difference between the NYY group and the Other group. Print out both the observed mean difference and the interval.

```{r}

```

(D) Use bootstrap resampling with 10,000 resamples to investigate the sampling distribution of the sample mean for each group. Use a single loop to generate both sets of resamples. Compute the mean, standard error, skewness, and kurtosis and create a normal probability plot for each of the two bootstrap distributions. Is either distribution approximately normal? Which has the smaller standard error and why?

```{r}

```

ANSWER:


### PROBLEM 3

In December 1969, the U.S. Selective Service System conducted a lottery to a determine draft order for eligible men in the upcoming year. The 366 days of the year (including February 29) were written on small pieces of paper. Each was placed in a plastic capsule, then the capsules were hand-mixed in a shoebox and dumped in a glass jar. Capsules were pulled out one at a time. The first number chosen was 258 (September 14). Anyone born on that day in 1944 through 1950 was assigned lottery number 1. The first 195 birthdates drawn were later called to serve in the order they were taken from the jar. Some people asserted that the  numbers were not randomly distributed through the year, in particular between the beginning and end of the year. (data: vietnamdraft.csv)

(A) Use `ggplot2` to create a density plot for the 1970 draft numbers. Then use `ggplot2` to create semi-transparent overlapping density plots to compare the 1970 draft numbers from the first half and second half of the year. 

```{r}

```

(B) Conduct a two-sample permutation test to determine whether the mean draft number in the first half of the year is different than the mean draft number in the second half of the year. What do you conclude? If they are different, which half had a smaller mean?

```{r}

```

ANSWER: 

(C) Conduct a two-sample permutation test to determine whether the proportion of draft numbers less than or equal to 195 was different in the first half of the year versus the second half of the year. What do you conclude? If they are different, which half had a larger proportion?

```{r}

```

ANSWER: 

(D) Based on your analyses, were people right to question the randomness of the system? Explain.

ANSWER: 


### PROBLEM 4

There are different rules of thumb for when it is appropriate to use a normal approximation to the binomial distribution. Fit depends on both the sample size and the value of p, since both affect the degree of symmetry. One of the least restrictive guidelines says it is appropriate to use the normal approximation when np &ge; 5, which means we theoretically could use a normal approximation for our Zener card test. Recall that the test is alleged to be statistically significant when someone gets 10 or more cards correct. 

(A) What is the probability that someone who is just guessing will get 10 or more correct, using exact Binomial calculation?

```{r}

```

(B) What is the probability that someone who is just guessing will get 10 or more correct, using the Normal approximation without continuity correction?

```{r}

```

(C) What is the probability that someone who is just guessing will get 10 or more correct, using the Normal approximation with continuity correction?

```{r}

```

(D) Does either of the Normal approximations seem reasonable? Explain.

ANSWER: 


<hr>

DONE!!!