---
title: "Exam 1"
author: "YOUR NAME HERE"
date: "`r Sys.time()`"
output: html_document
---

```{r, echo=FALSE, message=FALSE}
# Additional packages
library(car)
library(dplyr)
library(ggplot2)
library(moments)

# Oscar Winner Ages data
OSCAR <- read.csv(file=url("https://raw.githubusercontent.com/STAT-JET-ASU/Datafiles/master/Instructor/oscarages.csv"))

# Mauna Loa CO2 data
CO2 <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datafiles/master/Instructor/maunaloaCO2.csv")

# Anthropometric data
ANT <- read.csv(file=url("https://raw.githubusercontent.com/STAT-JET-ASU/Datafiles/master/Instructor/anthropometric.csv"))
```

**NOTE: Descriptive statistics should be computed using the `summarize()` function. All plots should have appropriate titles and axis labels. If you need to use a calculated numerical answer in a subsequent analysis, you should store it and call it back when needed, not type the value by hand. If you store a calculated answer, be sure to print it as well so I can see it.**

### PROBLEM 1

A digital communication channel transmits binary signals (digits 0 or 1). Suppose a particular system is susceptible to electronic interference, so there is a 5% chance that any bit that is sent will be incorrectly received (i.e., we send a 0 and it is mistakenly interpreted as 1). Assume that the interference is non-patterned, so all bit transmission errors are random and independent of one another.

One simple method for error reduction is repetition code. For critical messages, the digits are transmitted as repeated blocks of five identical digits. For example, a single 0 in the message would be sent as the block 00000 and a single 1 would be sent as 11111. The receiver uses "majority logic decoding" on blocks. If the majority of the five-block's digits are interpreted by the receiver as 0, then the transmitted digit represented by the block is recorded as a single 0. If the majority are interpreted as 1, then the transmitted digit is recorded as a 1.

A) What is the probability that any single five-block (e.g., 00000) is interpreted correctly?

```{r}

```

B) In the ASCII code, each letter is represented by a sequence of eight 0's and 1's. For example, 01001010 is the capital letter J. If we transmit a letter using the 5-block system described above, what is the probability that the letter will be interpreted correctly at the receiving end?

```{r}

```

C) The letters of Dr. Thomley's first name would be J = 01001010, i = 01101001, l = 01101100, and l = 01101100 in ASCII. What is your LAST name in ASCII? Modify the code below to find it.

```{r, message=FALSE}
require(gtools)
require(broman)
require(BMS)
NAMELET  <- c("J","i","l","l")
n        <- length(NAMELET)
NAMEHEX  <- convert2hex(asc(NAMELET))
NAMEBIN  <- matrix(rep(0,n*8),ncol=8)

for (i in 1:n) {NAMEBIN[i,1:8] <- hex2bin(NAMEHEX[i])}
rownames(NAMEBIN) <- NAMELET
colnames(NAMEBIN) <- c("Bit1","Bit2","Bit3","Bit4","Bit5","Bit6","Bit7","Bit8")
print(NAMEBIN)
```

D) What is the probability that your last name would be correctly transmitted?

```{r}

```

E) We plan to transmit the entire Gettysburg Address, which has 1,453 characters, including  letters and the spaces between words. A space is treated exactly like a letter and its code is 00100000. How many character errors should we expect to get? What is the standard deviation of the number of character errors? Solve using both theoretical calculations and simulation.

```{r}

```


### PROBLEM 2

In class, we used the Zener Card Tests case study on AsULearn to discuss psychic testing using Zener cards and how it can be modeled using the Binomial distribution. A standard Zener card deck has twenty-five cards, five each of five different symbols: circle, cross, waves, square, and star. The test on the web site we used has the person taking the test attempt to identify 25 Zener cards drawn with replacement from a standard deck (i.e., an "open" deck in their description). The test records the number of cards the person gets correct. (Note: The questions below may differ slightly from our class discussions.)

The test we used: http://www.psychic-experiences.com/psychic-tests/zener-cards.php

A) The Zener test web site says, "a score of 10 correct guesses or more out of 25 is required to prove some degree of psychic abilities." What is the probability that someone who is just guessing will get 10 or more correct?

```{r}

```

B) In statistical hypothesis testing, it is common to use the most extreme 5% of values as a cutoff for statistical significance. If we consider the upper 5% of the distribution as extreme in this case (i.e., a one-tailed test), how many correct guesses would a person have to exceed in the web site test to "prove" some degree of psychic ability?

```{r}

```

C) Suppose someone IS psychic? That person should be able to get more cards correct than someone who is just guessing, so what is their chance of getting 10 or more cards correct? Create a plot that shows x = chance of identifying a single card correctly (from guessing to 0.9, increasing in increments of 0.05) and y = probability of getting 10 or more cards correct out of 25.

```{r}

```

D) Dr. Thomley has a Zener card deck typically used by performing illusionists. It has 10 circles, 10 crosses, 10 waves, 10 squares, 10 stars, and one each of the following symbols: XX, spiral, null set, infinity, barbell, and triangle. Use appropriate functions to efficiently create a vector to represent this deck. Then draw a sample of size 10,000 from the deck and use the `table` function to summarize the results. Do the proportions of each symbol in your sample seem correct? Explain.

```{r}

```

ANSWER: 

E) If we use the deck in (D) instead of the standard Zener deck, can we model the number of correct guesses using a Binomial distribution? Why or Why not?

ANSWER: 


### PROBLEM 3

In January 2007, Entertainment Weekly magazine featured a cover story about three Academy Award nominees for Best Actress: Meryl Streep (age 57), Judy Densch (age 72), and Helen Mirren (age 61). A major focus of the article was the women’s ages and the perceived Hollywood bias in favor of younger actresses, a discussion that persists in one form or another to this day.

A) Create side-by-side boxplots to compare the ages of Best Actor, Best Actress, Best Supporting Actor, and Best Supporting actress winners prior to 2007.

```{r}

```

B) What is the upper cutoff for outliers for each of the four groups in (A)?

```{r}

```

B) Which actress won the Best Actress Oscar in 2007? How old was she? For what movie did she win? Determine this using code, not visual inspection of the data.

```{r}

```

C) Would the winner in (C) have been an outlier compared to previous Best Actress winners? Would  she have been an outlier compared to previous Best Actor winners, if she were a man of the same age? Determine this using code.

```{r}

```

D) Using the entire dataset, create a new variable to indicate which winners are classified as actresses and which are classified as actors. Then use this variable to construct semi-transparent, overlapping density plots to compare the age distributions of actresses and actors. 

```{r}

```

E) Compute the mean, standard deviation, minimum, and maximum for each group in (E).

```{r}

```

F) In what way could these analyses be used support an assertion of age bias againt women in Hollywood? What are the limitations of this sample if we want to generalize to the population of all actors and actresses?

ANSWER:


### PROBLEM 4

Carbon dioxide is one of the primary greenhouse gasses in earth's atmosphere. CO2 enters the atmosphere via burning fossil fuels, solid waste, trees and wood products, and certain chemical reactions. Plants absorb carbon dioxide from the atmopshere as part of the carbon cycle and some if it dissolves in the oceans. The Mauna Loa Observatory in Hawaii has been measuring carbon dioxide in earth's atmosphere since the 1950's. The observatory is 11,000 feet above sea level and its location makes it well-suited for gathering the kind of particulate data used to investigate climate change. The dataset includes raw CO2 values for each month (averaged across all days in the month) and adjusted CO2 values for the same months, where the adjustments remove the cyclic seasonal variation caused by plant growth cycles.

A) Create a plot that shows the values of both CO2 and adjusted CO2 over time for the entire dataset. 

```{r}

```

B) Compute means and standard deviations of adjusted monthly CO2 by decade. Exclude the incomplete decades 1950s and 2010s. Graph the means on one barplot and the standard deviations on another.

```{r}

```

C) Given the patterns you observe, estimate what you think the adjusted CO2 mean and standard deviation might be for the 2010s, once the decade is over. Discuss the rationale for your estimation method.

```{r}

```

ANSWER: 

D) How could these data be used to support the assertion that global climate change is occurring? Why are data such as these evidence rather than proof?

ANSWER:


### PROBLEM 5

In class, we examined the Height variable in the Anthropometric dataset and discussed that the distribution appears to be multimodal. It may be that males and females have different height distributions, so it doesn't make sense to mix them in the analysis. If we analyze self-identified males and females separately, are the distributions approximately normal? Do they have similar centers? What about spreads? Use appropriate graphs and summaries to make this determination.

```{r}

```

ANSWER: 

<hr>

DONE!!!