---
title: "Introduction to<br>Sampling and the CLT"
author: "Author: Jill E. Thomley"
date: '`r format(Sys.time(), "%B %d, %Y @ %I:%M %p")`'
output: 
  ioslides_presentation:
    logo: images/logoASU.jpg
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, comment = "")
library(tidyverse)
```


## Before We Begin...

These slides are not meant to be standalone information. You should take notes to flesh out the contents. I recommend that you create an R Markdown document where you can combine information and code from the slides and your own additional notes and explorations to make connections.

**Related Materials**

* Ch XXX of *Mathematical Statistics with Resampling and R, 2^nd^ Ed.*
* Ch 7 of [*Modern Dive*](https://moderndive.com/)
* DataCamp [Introduction to R](https://www.datacamp.com/courses/free-introduction-to-r)
* DataCamp [Foundations of Probability in R](https://www.datacamp.com/courses/foundations-of-probability-in-r)

Refer to "[4.1 ioslides presentation](https://bookdown.org/yihui/rmarkdown/ioslides-presentation.html)" in [*R Markdown: The Definitive Guide*](https://bookdown.org/yihui/rmarkdown/) if you are interested in creating R Markdown slides.


## 

<iframe width="560" height="315" src="https://www.youtube.com/embed/Zbw-YvELsaM" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Video Sampling Example in R

Create the population and calculate the population mean $\mu$.

```{r}
ages <- c(234,241, 233, 227, 251, 227, 242, 239, 241, 238, 230, 246, 
          231, 243, 238, 276)
mu <- mean(ages)
```

Set up a loop to take repeated samples. Calculate the mean of each sample and store it in a vector. Here, `sims` is the number of samples we plan to take.

```{r}
sims <- 10^5
samplemeans <- numeric(sims)
for (i in 1:sims) {
  samplemeans[i] <- mean(sample(ages, 3))
}
```


## 

```{r}
ggplot(NULL, aes(x = samplemeans)) +
  geom_histogram(bins = 15, fill = "lightblue", color = "gray") +
  geom_vline(xintercept = mu, color = "red") + theme_linedraw()
```


##

We selected three people from the class (without replacement) for each sample. How many unique samples are possible? This can be determined with combinatorics. Let $N$ be the population size and $n$ be the sample size. 

$$N \text{ choose } n = {N \choose n} = \frac{N!}{n!(N-n)!}$$

$${16 \choose 3} = \frac{16!}{3!(16-3)!} = \frac{16 \times 15 \times 14 }{3 \times 2 \times 1} = \frac{3360}{6} = 560$$

We can calculate this answer using the `choose()` function in `R`.

```{r}
choose(16, 3)
```


## 


##


## 

<iframe width="560" height="315" src="https://www.youtube.com/embed/q50GpTdFYyI" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>


## 

<iframe width="560" height="315" src="https://www.youtube.com/embed/Pujol1yC1_A" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Exponential Simulation Using R




## Video Salary Example Using R

Salaries at a very large corporation have a mean of $62,000 and a standard deviation of $32,000. If a single employee is selected, what is the probability that their salary exceeds $66,000?

```{r}
pnorm(66000, mean = 62000, sd = 32000, lower.tail = FALSE)
```

If 100 employees are selected, what is the probability that their salary exceeds $66,000?

```{r}
pnorm(66000, mean = 62000, sd = 32000/sqrt(100), lower.tail = FALSE)
```

