---
title: "Estimation Example: Bangladesh Wells (proportion)"
author: "Created by Jill E. Thomley"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: TRUE
    toc_depth: 4
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

### **Packages Used**

```{r}
library(resampledata)
library(tidyverse)
library(moments)
library(infer)
library(knitr)
library(kableExtra)
```

### **Problem & DataSet**

### **Bangladesh Well Data**

Bangladesh has suffered the largest population-level poisoning in history. Millions of people were exposed to naturally-occurring arsenic from tube-wells. Ironically, the wells had been installed to provide cleaner water than surface sources in order to reduce gastrointestinal diseases. Studies in other countries with long-term exposure to arsenic in groundwater found that 1 in 10 people who consume water containing 500 micrograms of arsenic per liter (mcg/l) may ultimately die from cancers caused by arsenic. The `Bangladesh` dataset from the textbook *Mathematical Statistics with Resampling and R* (found in the `resampledata` package) contains a sample of tested wells.

**Problem #1:** What is the true proportion $p$ of Bangladesh wells that exceed 500 micrograms of arsenic per liter?

**Problem #2:** What is the true proportion $p$ of wells that exceed the WHO standard of 10 micrograms of arsenic per liter?

```{r}
glimpse(Bangladesh)
```

### **Exploratory Data Analysis #1**

**Problem #1:** What proportion of Bangladesh wells exceed 500 micrograms of arsenic per liter?

#### **Summary Statistics**

The sample proportion $\hat{p}$ is our point estimate for the true proportion $p$.

```{r}
Bangladesh <- Bangladesh %>% 
  mutate(Over500 = ifelse(Arsenic > 500, "Does Exceed 500", "Does Not Exceed 500"))

summaries <- Bangladesh %>% 
  summarize(n = n(),
            p_hat = mean(Over500 == "Does Exceed 500"))
summaries
```

#### **Barplot of Proportions**

```{r}
ggplot(Bangladesh, aes(x = Over500)) +
  geom_bar((aes(y = (..count..)/sum(..count..))), color = "gray", fill = "skyblue") +
  labs(title = "Arsenic Contamination of Tube Wells in Balgladesh",
       x = "Does the arsenic concentration exceed 500 mcg/l?",
       y = "proportion of wells") +
  theme_linedraw()
```

### **Confidence Interval Estimation #1**

**Problem #1:** What proportion of Bangladesh wells exceed 500 micrograms of arsenic per liter?

#### **Conduct Bootstrap Simulation**

Use bootstrap simulation to estimate the sampling distribution of $\hat{p}$.

```{r}
bootstrap_distribution <- Bangladesh %>% 
  specify(response = Over500, success = "Does Exceed 500") %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "prop")
```

#### **Plot the Results and Assess**

```{r}
visualize(bootstrap_distribution) 

ggplot(bootstrap_distribution, aes(sample = stat)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Simulation-Based Bootstrap Distribution")

bootstrap_distribution %>% 
  summarize(skew = skewness(stat),
            kurt = kurtosis(stat),
            kurt_adj = kurt - 3)
```

The shape of the sampling distribution of $\hat{p}$ looks mildly right-skewed on the plots, but the skewness (which is less than 0.5) and the kurtosis (approximately 3) suggest that it is not far from approximately normal.

#### **95% Percentile Interval for $p$**

Find the middle 95% of the bootstrap distribution to create an interval estimate.

```{r}
percentile_ci <- bootstrap_distribution %>% 
  get_confidence_interval(level = 0.95, type = "percentile")

percentile_ci %>% 
  kable(digits = 4, align = "ccc") %>% 
  kable_styling(full_width = FALSE, position = "left", font_size = 14)
```

#### **95% Confidence Interval Using SE**

Use the "plug in" method $\hat{p} \pm 1.96 \times SE$ to create an interval estimate. The original sample proportion (above) is $\hat{p}$ and $SE$ is the standard deviation of the generated bootstrap distribution. The sampling distribution of $\hat{p}$ must be approximately normal for this method to be valid.  

```{r}
standard_error_ci <- bootstrap_distribution %>% 
  get_confidence_interval(level = 0.95, type = "se", point_estimate = summaries$p_hat)

standard_error_ci %>% 
  kable(digits = 4, align = "ccc") %>% 
  kable_styling(full_width = FALSE, position = "left", font_size = 14)
```

#### **Traditional 95% Confidence Interval**

The traditional *z*-based confidence interval assumes that we have an independent random sample and a large enough sample size that the sampling distribution of $\hat{p}$ is approximately normal. The plots for the bootstrap distribution suggests that the latter is true.

The `prop.test()` finds the CI for the category that comes first alphabetically, so you may have to recode the categories to get the estimate you want. 

```{r}
prop.test(table(Bangladesh$Over500))$conf.int
```

#### **Comparing the Estimates**

<details>
```{r}
comparison <- rbind(percentile_ci, 
      standard_error_ci,
      tibble(lower_ci = prop.test(table(Bangladesh$Over500))$conf.int[1],
             upper_ci = prop.test(table(Bangladesh$Over500))$conf.int[2])) %>% 
  mutate(type = c("Percentile", "SE Method", "Traditional")) %>% 
  select(type, lower_ci, upper_ci) %>% 
  kable(digits = 4, align = "lcc") %>% 
  kable_styling(full_width = FALSE, position = "left", font_size = 14)

comparison_plot <- visualize(bootstrap_distribution) +
  shade_ci(endpoints = percentile_ci, color = "skyblue", fill = "white") +
  shade_ci(endpoints = standard_error_ci, color = "orange", fill = "white") +
  labs(subtitle = "blue = 95% percentile interval, orange = 95% SE interval")
```
</details>

```{r}
comparison
```

<hr>

### **Exploratory Data Analysis #2**

**Problem #2:** What proportion of Bangladesh wells exceed 10 micrograms of arsenic per liter?

#### **Summary Statistics**

The sample proportion $\hat{p}$ is our point estimate for the true proportion $p$.

```{r}
Bangladesh <- Bangladesh %>% 
  mutate(Over10 = ifelse(Arsenic > 10, "Does Exceed 10", "Does Not Exceed 10"))

summaries <- Bangladesh %>% 
  summarize(n = n(),
            p_hat = mean(Over10 == "Does Exceed 10"))
summaries
```

#### **Barplot of Proportions**

```{r}
ggplot(Bangladesh, aes(x = Over10)) +
  geom_bar((aes(y = (..count..)/sum(..count..))), color = "gray", fill = "skyblue") +
  labs(title = "Arsenic Contamination of Tube Wells in Balgladesh",
       x = "Does the arsenic concentration exceed 500 mcg/l?",
       y = "proportion of wells") +
  theme_linedraw()
```

### **Confidence Interval Estimation #2**

**Problem #2:** What proportion of Bangladesh wells exceed 10 micrograms of arsenic per liter?

#### **Conduct Bootstrap Simulation**

Use bootstrap simulation to estimate the sampling distribution of $\hat{p}$.

```{r}
bootstrap_distribution <- Bangladesh %>% 
  specify(response = Over10, success = "Does Exceed 10") %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "prop")
```

#### **Plot the Results and Assess**

```{r}
visualize(bootstrap_distribution) 

ggplot(bootstrap_distribution, aes(sample = stat)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Simulation-Based Bootstrap Distribution")

bootstrap_distribution %>% 
  summarize(skew = skewness(stat),
            kurt = kurtosis(stat),
            kurt_adj = kurt - 3)
```

The shape of the sampling distribution of $\hat{p}$ looks approximately normal.

#### **95% Percentile Interval for $p$**

Find the middle 95% of the bootstrap distribution to create an interval estimate.

```{r}
percentile_ci <- bootstrap_distribution %>% 
  get_confidence_interval(level = 0.95, type = "percentile")

percentile_ci %>% 
  kable(digits = 4, align = "ccc") %>% 
  kable_styling(full_width = FALSE, position = "left", font_size = 14)
```

#### **95% Confidence Interval Using SE**

Use the "plug in" method $\hat{p} \pm 1.96 \times SE$ to create an interval estimate. The original sample proportion (above) is $\hat{p}$ and $SE$ is the standard deviation of the generated bootstrap distribution. The sampling distribution of $\hat{p}$ must be approximately normal for this method to be valid.  

```{r}
standard_error_ci <- bootstrap_distribution %>% 
  get_confidence_interval(level = 0.95, type = "se", point_estimate = summaries$p_hat)

standard_error_ci %>% 
  kable(digits = 4, align = "ccc") %>% 
  kable_styling(full_width = FALSE, position = "left", font_size = 14)
```

#### **Traditional 95% Confidence Interval**

The traditional *z*-based confidence interval assumes that we have an independent random sample and a large enough sample size that the sampling distribution of $\hat{p}$ is approximately normal. The plots for the bootstrap distribution suggests that the latter is true.

The `prop.test()` finds the CI for the category that comes first alphabetically, so you may have to recode the categories to get the estimate you want. 

```{r}
prop.test(table(Bangladesh$Over10))$conf.int
```

#### **Comparing the Estimates**

<details>
```{r}
comparison <- rbind(percentile_ci, 
      standard_error_ci,
      tibble(lower_ci = prop.test(table(Bangladesh$Over10))$conf.int[1],
             upper_ci = prop.test(table(Bangladesh$Over10))$conf.int[2])) %>% 
  mutate(type = c("Percentile", "SE Method", "Traditional")) %>% 
  select(type, lower_ci, upper_ci) %>% 
  kable(digits = 4, align = "lcc") %>% 
  kable_styling(full_width = FALSE, position = "left", font_size = 14)

comparison_plot <- visualize(bootstrap_distribution) +
  shade_ci(endpoints = percentile_ci, color = "skyblue", fill = "white") +
  shade_ci(endpoints = standard_error_ci, color = "orange", fill = "white") +
  labs(subtitle = "blue = 95% percentile interval, orange = 95% SE interval")
```
</details>

```{r}
comparison
```

All the values in the confidence interval(s) are above 0.50, so we can be 95% confident that more than half the wells exceed the WHO standard.


<hr>
```{r}
sessionInfo()
```

