---
title: 'HW: Foundations of Inference in `R`'
author: 'YOUR NAME HERE'
date: '`r format(Sys.time(), "%B %d, %Y @ %I:%M %p")`'
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: false
    theme: yeti
    highlight: textmate
---

**DataCamp Course Link:** [Foundations of Inference](https://www.datacamp.com/courses/foundations-of-inference)

*One of the foundational aspects of statistical analysis is inference, or the process of drawing conclusions about a larger population from a sample of data. Although counter intuitive, the standard practice is to attempt to disprove a research claim that is not of interest. For example, to show that one medical treatment is better than another, we can assume that the two treatments lead to equal survival rates only to then be disproved by the data. Additionally, we introduce the idea of a p-value, or the degree of disagreement between the data and the hypothesis. We also dive into confidence intervals, which measure the magnitude of the effect of interest (e.g. how much better one treatment is than another).*

***

**Instructions**

> Replicate all of the activities from the DataCamp course _Foundations of Inference_, including the code and comments. For videos, summarize the main points. You are welcome to include additional text or comments of your own to help you recall the specifics of the lessons.

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

### Packages Used

```{r packages}
library(tidyverse)
library(infer)
library(NHANES) #includes NHANES dataset
```

### Datasets Used

```{r datasets}
disc <- readRDS(gzcon(url("https://assets.datacamp.com/production/repositories/538/datasets/60566129b391ef827ea9c8a9846608dee24ce34a/disc_new.rds")))
str(disc)

disc_small <- readRDS(gzcon(url("https://assets.datacamp.com/production/repositories/538/datasets/543fa990550c61f6a2cc175b0a0414528f8094c0/disc_small.rds")))
str(disc_small)
  
disc_big <- readRDS(gzcon(url("https://assets.datacamp.com/production/repositories/538/datasets/f03da8fc4a2ae50a3ddf775324f4df90c96f7f26/disc_big.rds")))
str(disc_big)

opportunity <- tibble(decision = rep(c("buyDVD", "nobuyDVD"), c(97, 53)),
                      group = rep(rep(c("control", "treatment"), 2), c(56, 41, 19, 34)))
str(opportunity)

load(url("https://assets.datacamp.com/production/repositories/538/datasets/b1071ca5cb72143820e33fd7c6605dc4b3f11b7a/all_polls.RData"))
str(all_polls)
```

***
### <span style="font-size: 100%; color: #2780E3;">&#9312;</span> Introduction to Ideas of Inference

#### Welcome to the course!

VIDEO SUMMARY:

#### Hypothesis (1)

NULL HYPOTHESIS:

#### Hypothesis (2)

ALTERNATIVE HYPOTHESIS:

#### Randomized distributions

VIDEO SUMMARY:

#### Working with the NHANES data

**Note:** You do not need to reload `NHANES` and `ggplot` here. 

```{r}
# Exercise 1

```

```{r}
# Exercise 2

```

```{r}
# Exercise 3

```

#### Calculating the statistic of interest

```{r}
# Exercise 1

```

```{r}
# Exercise 2

```

#### Randomized data under null model of independence

```{r}
# Exercise 1

```

```{r}
# Exercise 2

```

```{r}
# Exercise 3

```

#### Randomized statistics and dotplot

```{r}
# Exercise 1

```

```{r}
# Exercise 2

```

#### Randomization density

```{r}

```

#### Using the randomized distribution

VIDEO SUMMARY:

#### Do the data come from the population?

```{r}

```

#### What can you conclude?

CONCLUSION:

#### The study conclusions

VIDEO SUMMARY:

***
### <span style="font-size: 100%; color: #2780E3;">&#9313;</span> Completing a Randomization Test: Gender Discrimination

#### Example: gender discrimination

VIDEO SUMMARY:

#### Example: Gender discrimination hypotheses

$H_0:$

$H_A:$

#### Summarizing gender discrimination

```{r}

```

#### Step-by-step through the permutation

```{r}

```

#### Randomizing gender discrimination

```{r}
# Exercise 1

```

```{r}
# Exercise 2

```

```{r}
# Exercise 3

```

#### Distribution of statistics

VIDEO SUMMARY:

#### Reflecting on analysis

```{r}

```

#### Critical region

```{r}

```

#### Two-sided critical region

```{r}

```

#### Why 0.05?

VIDEO SUMMARY:

#### How does the sample size affect results?

CONCLUSION: The statistic of 0.2917 would ...

#### Sample size in randomization distribution

```{r}

```

#### Sample size for critical region

```{r}

```

#### What is a p-value?

VIDEO SUMMARY:

#### Calculating the p-values

```{r}

```

#### Practice calculating p-values

```{r}
# Exercise 1

```

```{r}
# Exercise 2

```

```{r}
# Exercise 3

```

#### Calculating two-sided p-values

```{r}

```

#### Summary of gender discrimination

VIDEO SUMMARY:

***
### <span style="font-size: 100%; color: #2780E3;">&#9314;</span> Hypothesis Testing Errors: Opportunity Cost

#### Example: opportunity cost

VIDEO SUMMARY:

#### Summarizing opportunity cost (1)

```{r}

```

#### Plotting opportunity cost

```{r}

```

#### Randomizing opportunity cost

```{r}
# Exercise 1

```

```{r}
# Exercise 2

```

```{r}
# Exercise 3

```

#### Summarizing opportunity cost (2)

```{r}

```

#### Opportunity cost conclusion

CONCLUSION:

#### Errors and their consequences

VIDEO SUMMARY:

#### Different choice of error rate

DECISION:

#### Errors for two-sided hypotheses

Type I Error:

Type II Error:

#### p-value for two-sided hypotheses: opportunity costs

```{r}

```

#### Summary of opportunity costs

VIDEO SUMMARY:

***
### <span style="font-size: 100%; color: #2780E3;">&#9315;</span> Confidence Intervals

#### Parameters and confidence intervals

VIDEO SUMMARY:

#### What is the parameter?

PARAMETER: 

#### Hypothesis test or confidence interval?

CHOICE:

#### Bootstrapping

VIDEO SUMMARY:

#### Resampling from a sample

```{r}
# Exercise 1

```

```{r}
# Exercise 2

```

```{r}
# Exercise 3

```

#### Visualizing the variability of p-hat

```{r}

```

#### Always resample the original number of observations

HOW MANY OBSERVATIONS:

#### Variability in p-hat

VIDEO SUMMARY:

#### Empirical Rule

```{r}

```

#### Bootstrap t-confidence interval

```{r}

```

#### Bootstrap percentile interval

```{r}

```

#### Interpreting CIs and technical conditions

VIDEO SUMMARY:

#### Sample size effects on bootstrap CIs

```{r}

```

#### Sample proportion value effects on bootstrap CIs

```{r}

```

#### Percentile effects on bootstrap CIs

```{r}

```

#### Summary of statistical inference

VIDEO SUMMARY:

***
```{r}
sessionInfo()
```
