---
title: "Last Day Activity"
author: "YOUR NAME"
date: "6/27/2018"
output: html_document
---

### PROBLEM 1

In class, we worked problems 6-10 in the Sampling Distributions Part 1 practice quiz. Verify your answers using simulation.

```{r}

```

### PROBLEM 2

Consider the Verizon repair times dataset in the book. For the CLEC group...

* find the sample mean, sample standard deviation, and smaple size
* generate a bootstrap sampling distribution for x-bar using 10,000 resamples
* compute the mean and standard deviation of the bootstrap distribution
* compute a 95% percentile interval to estimate the population mean of repair times
* compute a classical t-based 95% confidence interval 

Are your interval results the same? Is the t-based approach appropriate here? Why or why not? (Note: time is in units of hours.)

So you can check your work:

sample mean: 16.509
sample sd: 19.504
sample size: 23
bootstrap mean: about 16.509
bootstrap sd: about 19.504 / sqrt(23)
bootstrap interval: about (10 to 25)
t-based interval: (8.075 to 24.943)
We need CLT to be true to use t-based approach...

```{r}
VZ <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/Verizon.csv")

VZ2 <- VZ %>% filter(Group == "CLEC")

mean(VZ2$Time)
sd(VZ2$Time)
length(VZ2$Time)

N <- 10^4
bootmeans <- numeric(N)
for (i in 1:N){
  bootsample   <- sample(VZ2$Time, replace = TRUE)
  bootmeans[i] <- mean(bootsample)
  }

mean(bootmeans)
sd(bootmeans)
quantile(bootmeans, c(0.025, 0.975))
hist(bootmeans)
require(car)
qqPlot(bootmeans)

VZ2 <- VZ %>%  filter(Group == "CLEC")
mean(VZ2$Time) - qt(0.975, length(VZ2$Time) - 1) * sd(VZ2$Time) / sqrt(length(VZ2$Time))
mean(VZ2$Time) + qt(0.975, length(VZ2$Time) - 1) * sd(VZ2$Time) / sqrt(length(VZ2$Time))
```

