---
title: "Case Study: Anthropometric Data"
author: "PUT YOUR NAME HERE"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: TRUE
    toc_depth: 4
---

<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

### Load the Package(s)

```{r loadpackages}
library(tidyverse)
library(moments) # for skewness and kurtosis
library("gridExtra") # for bonus end content
library(grid)        # for bonus end content
```

### Problem Description

In our [Modeling & Regression: Theory and Line Fitting](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/ModelingRegression01.html#1) slides, we posed the question of whether people are squares, as asserted by the ancient Roman architect Vitruvius. In other words, are people's heights and armspans equal? Refer to the [description page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/anthropometric.html) for Dr. Thomley's [anthropometric dataset](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/anthropometric.csv) for context regarding the data we will use to assess this claim. These data were collected over several semesters from students in Dr. Thomley’s STT1810 and STT3850 courses. Students measured one another using tape measures and/or yardsticks. Gender was self-identified.

Before doing any modeling, it is good practice to conduct exploratory data analysis on the variables in order to assess shape and determine whether there are any outliers or other anomalies, which might affect our models.

Use `tidyverse` methods (e.g., `dplyr` and `ggplot`) as much as possible when answering the questions. 

### Import the Dataset

Use `read_csv` to import the data, then `glimpse` and `head` to display the dataset structure.

```{r loaddatasets}
anthro <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/anthropometric.csv"))
glimpse(anthro)
head(anthro)
```

### Exploratory Analysis 1

#### What genders are present in the data?

Create a frequency table and barplot for the variable `gender`. 

```{r}
anthro %>% 
  count(gender)
```

For exploration, we can do a basic barplot. The heights of the bars should match the table. Notice there are very few people (so far) who have self-identified anything besides `male` or `female`.

```{r}
ggplot(anthro, aes(x = gender)) +
  geom_bar()
```

For presentation, we should give the plot an informative title and axis labels. Adding color is optional, but it can be nice to have. Try to keep colorblind-safe options in mind if you are using colors to convey information.

```{r}
ggplot(anthro, aes(x = gender)) +
  geom_bar(fill = "purple") + 
  labs(title = "Self-Reported Student Gender", 
       subtitle = "From several of Dr. Thomley's STT1810 and STT3850 courses",
       x = "self-identified gender",
       caption = "source: anthropometric.csv dataset")
```

#### What semesters are present in the data?

Create a frequency table and barplot for the variable `semester`. 

The code is the same as what we used above, just change the variable. Notice that the semesters are in alphabetical order, which is `R`'s default for character vectors.

```{r}
anthro %>% 
  count(semester)

# with some example bar formatting
ggplot(anthro, aes(x = semester)) +
  geom_bar(fill = "orange", color = "orangered", width = 0.75)
```

#### Putting the semesters in correct order

Make `semester` a `factor` with the categories in the correct order. Then make frequency table and barplot using this new variable.

Academic years go spring, summer, fall. Put the years in ascending order.

```{r}
sem_levels <- c("S13", "Sum13", "F13", "S14", "Sum14", "Sum15", "F15", "S16", "Sum16", "S17", "S18")

anthro <- anthro %>% 
  mutate(semester = factor(semester, levels = sem_levels))

anthro %>% 
  count(semester)

ggplot(anthro, aes(x = semester)) +
  geom_bar()
```

### Exploratory Analysis 2

#### Plots and summaries for `height`

Create a histogram, density plot, and boxplot for the variable `height`. Also calculate skewness, kurtosis, and the upper and lower fences.

This shows several `bin` options for the histogram. You often have to try several options to get a good representation. These data seem to have more than one mode (peak or local maxima)---either two or three, judging by the histograms and density plot. We may need to explore the data in groups to uncover what is going on.

```{r}
ggplot(anthro, aes(x = height)) +
  geom_histogram()   # default number of bins is about 30

ggplot(anthro, aes(x = height)) +
  geom_histogram(bins = 25)   # we will get about 25 bins

ggplot(anthro, aes(x = height)) +
  geom_histogram(bins = 15)   # we will get about 15 bins

ggplot(anthro, aes(x = height)) +
  geom_histogram(bins = 10)   # we will get about 10 bins

ggplot(anthro, aes(x = height)) +
  geom_histogram(binwidth = 10)  # bins are 10 units wide

ggplot(anthro, aes(x = height)) +
  geom_histogram(binwidth = 2)    # bins are 2 units wide
```

We can also add color and labels to histograms.

```{r}
ggplot(anthro, aes(x = height)) +
  geom_histogram(bins = 12, fill = "skyblue", color = "blue") +
  labs(title = "Measured Heights of Students", 
       subtitle = "From several of Dr. Thomley's STT1810 and STT3850 courses",
       x = "height in inches",
       caption = "source: anthropometric.csv dataset")
```

The other plots follow the same pattern. The variable the same (same `aes`), just change the type of plot (different `geom`).

```{r}
ggplot(anthro, aes(x = height)) +
  geom_density()

ggplot(anthro, aes(x = height)) +
  geom_boxplot()
```

We can calculate all four statistics in one `summarize` statement. Refer to the entries for skewness and kurtosis in the glossary for more information. There is missing data, so we need to use `na.rm = TRUE` to ignore these `NA` values.

```{r}
anthro %>% 
  summarize(skew = skewness(height, na.rm = TRUE),
            kurt = kurtosis(height, na.rm = TRUE),
            LF   = fivenum(height)[2] - 1.5 * IQR(height, na.rm = TRUE),
            UF   = fivenum(height)[4] + 1.5 * IQR(height, na.rm = TRUE))
```

#### Plots and summaries for `armspan`

Create a histogram, density plot, and boxplot for the variable `armspan`. Also calculate skewness, kurtosis, and the upper and lower fences for the boxplot.

```{r}
ggplot(anthro, aes(x = armspan)) +
  geom_histogram(bins = 15)

ggplot(anthro, aes(x = armspan)) +
  geom_density()

ggplot(anthro, aes(x = armspan)) +
  geom_boxplot()

anthro %>% 
  summarize(skew = skewness(height, na.rm = TRUE),
            kurt = kurtosis(height, na.rm = TRUE),
            LF   = fivenum(height)[2] - IQR(height, na.rm = TRUE),
            UF   = fivenum(height)[4] + IQR(height, na.rm = TRUE))
```

#### Plots and summaries for `ideal`

Create a histogram, density plot, and boxplot for the variable `ideal` height. Also calculate skewness, kurtosis, and the upper and lower fences for the boxplot.

```{r}
ggplot(anthro, aes(x = ideal)) +
  geom_histogram(bins = 20)

ggplot(anthro, aes(x = ideal)) +
  geom_density()

ggplot(anthro, aes(x = ideal)) +
  geom_boxplot()

anthro %>% 
  summarize(skew = skewness(height, na.rm = TRUE),
            kurt = kurtosis(height, na.rm = TRUE),
            LF   = fivenum(height)[2] - IQR(height, na.rm = TRUE),
            UF   = fivenum(height)[4] + IQR(height, na.rm = TRUE))
```

### Exploratory Analysis 3

#### Filter the datasets for `gender`

Filter the data so it contains only males and females, since all other categories have too few observations for meaningful analysis. Hint: Use a filter with `%in%` to get the groups you want.

```{r}
anthroMFonly <- anthro %>% 
  filter(gender %in% c("male", "female"))
```

#### Faceted histogram

Create a histogram for `height` faceted by `gender`, with the two plots arranged one above the other (so they will share a common x-axis scaling  the final plot). Do the same for `armspan`.

Since histograms already have two quantitative axes (the variable on `x` and the count on `y`), we need to `facet` if we want to compare groups. Faceting is similar to a `group_by()` for plots. The `facet_grid()` lets us specify a matrix or "grid" arrangement for plots using a row variable and a column variable, as `facet_grid(rowvar ~ colvar)`. Using `gender` as the row variable means the plots will have the same x-axis and same bins. We only want to facet by one variable, so we replace the other with a `.`

Here I added some axis scaling so we can see the measurements more clearly. The `breaks` are the locations of the axis tick marks. Notice what I call the "six foot bump" at about 72" in the male graph. Where do you think that comes from?

```{r}
ggplot(anthroMFonly, aes(x = height)) +
  geom_histogram(bins = 18, fill = "skyblue", color = "blue") +
  scale_x_continuous(breaks = seq(from = 0, to = 100, by = 1)) + # tick marks
  facet_grid(gender ~ .) +
  labs(title = "Measured Heights of Students", 
       subtitle = "From several of Dr. Thomley's STT1810 and STT3850 courses",
       x = "height in inches",
       y = "self-identified gender",
       caption = "source: anthropometric.csv dataset")
```

#### Faceted density plot

Create a density plot for `ideal` faceted by `gender`, with the two plots arranged one above the other (so they will share a common x-axis scaling in the final plot).

Notice that the peaks are in different places on the x-axis for female and male distributions. This is why the combined distribution you explored above is bimodal.

```{r}
ggplot(anthroMFonly, aes(x = ideal)) +
  geom_density() +
  scale_x_continuous(breaks = seq(from = 0, to = 100, by = 2)) +
  facet_grid(gender ~ .) +
  labs(title = "Ideal Heights of Students", 
       subtitle = "From several of Dr. Thomley's STT1810 and STT3850 courses",
       x = "self-selected ideal height in inches",
       y = "self-identified gender",
       caption = "source: anthropometric.csv dataset") +
  theme_bw() # changes thematic elements like bg color
# https://ggplot2.tidyverse.org/reference/ggtheme.html
```

We could also layer the density plots, for a different view.

```{r}
ggplot(anthroMFonly, aes(x = ideal, fill = gender)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(breaks = seq(from = 0, to = 100, by = 2)) +
  labs(title = "Ideal Heights of Students", 
       subtitle = "From several of Dr. Thomley's STT1810 and STT3850 courses",
       x = "self-selected ideal height in inches",
       caption = "source: anthropometric.csv dataset")
```

#### Comparative boxplot

Create comparative boxplots for `height` split by `gender`. Also calculate the upper and lower fences for each gender. Do the same for `ideal` height.

Boxplots only have one quantitative axis, so we can group by specifying a categorical variable for the other axis. Boxplots may be horizontal or vertical. Making them horizontal keeps the x-axis consistent with the other plots.

```{r}
# horizontal
ggplot(anthroMFonly, aes(x = height, y = gender)) +
  geom_boxplot(fill = "skyblue") +
  labs(title = "Measured Heights of Students", 
       subtitle = "From several of Dr. Thomley's STT1810 and STT3850 courses",
       x = "height in inches",
       y = "self-identified gender",
       caption = "source: anthropometric.csv dataset")

# vertical
ggplot(anthroMFonly, aes(x = gender, y = height)) +
  geom_boxplot(fill = "orange") +
  labs(title = "Measured Heights of Students", 
       subtitle = "From several of Dr. Thomley's STT1810 and STT3850 courses",
       x = "self-identified gender",
       y = "height in inches",
       caption = "source: anthropometric.csv dataset")
```

If we want the plots to be different colors, we need to `fill` with our grouping variable in the aesthetics rather than specify a color in the `geom`. Then we can choose our colors with `scale_fill_manual()` or a similar fill function such as `scale_fill_brewer()`.

```{r}
ggplot(anthroMFonly, aes(x = height, y = gender, fill = gender)) +
  geom_boxplot() +
  labs(title = "Measured Heights of Students", 
       subtitle = "From several of Dr. Thomley's STT1810 and STT3850 courses",
       x = "height in inches",
       y = "self-identified gender",
       caption = "source: anthropometric.csv dataset") +
  scale_fill_manual(values = c("skyblue", "orange")) +
  theme(legend.position = "none") # removes the redundant legend
```

### Concluding Discussions

*I will leave you to think a bit more about these for now...*

#### Overall distribution shapes 

What are the overall shapes for the distributions of `height`, `armspan`, and `ideal`? Are they all unimodal? Are there any outliers?

#### Comparing male and female

Are there differences in distribution between males and females for `height`, `armspan`, and `ideal` height in this sample? Describe those differences.

#### Representativeness

What makes this sample *not* representative of people as a whole?

### BONUS: Using `grid.arrange()`

The `grid.arrange()` function allows us to group multiple plots into a useful or pleasing display of our choice. For several more options, see http://www.sthda.com/english/wiki/wiki.php?id_contents=7930.

```{r}
# histogram
plotA <- ggplot(anthro, aes(x = ideal)) +
  geom_histogram(bins = 20) +
  labs(x = "ideal height in inches")

# density
plotB <- ggplot(anthro, aes(x = ideal)) +
  geom_density() +
  labs(x = "ideal height in inches")

# horizontal boxplot
plotC <- ggplot(anthro, aes(x = ideal)) +
  geom_boxplot() +
  labs(x = "ideal height in inches",
       y = "")   # helps with spacing here

# ECDF plot
plotD <- ggplot(anthro, aes(x = ideal)) +
  stat_ecdf() +
  labs(x = "ideal height in inches")

# vertical boxplot
plotE <- ggplot(anthro, aes(y = ideal)) +
  geom_boxplot() +
  labs(y = "ideal height in inches")

grid.arrange(plotA, plotB, plotC, plotD, nrow = 2, ncol = 2,
             top = textGrob("Graphical Analysis of Ideal Height"))

grid.arrange(plotA, plotB, plotC, plotD, nrow = 4,
             top = textGrob("Graphical Analysis of Ideal Height"))

grid.arrange(plotE, arrangeGrob(plotA, plotB, plotD), ncol = 2,
             top = textGrob("Graphical Analysis of Ideal Height"))
```

```{r}
sessionInfo()
```

