---
title: "Case Study: Anthropometric Data"
author: "PUT YOUR NAME HERE"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: TRUE
    toc_depth: 4
---

<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

### Load the Package(s)

```{r loadpackages}
# Put your code here.
```

### Problem Description

In our [Modeling & Regression: Theory and Line Fitting](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/ModelingRegression01.html#1) slides, we posed the question of whether people are squares, as asserted by the ancient Roman architect Vitruvius. In other words, are people's heights and armspans equal? Refer to the [description page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/anthropometric.html) for Dr. Thomley's [anthropometric dataset](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/anthropometric.csv) for context regarding the data we will use to assess this claim. These data were collected over several semesters from students in Dr. Thomley’s STT1810 and STT3850 courses. Students measured one another using tape measures and/or yardsticks. Gender was self-identified.

Before doing any modeling, it is good practice to conduct exploratory data analysis on the variables in order to assess shape and determine whether there are any outliers or other anomalies, which might affect our models.

Use `tidyverse` methods (e.g., `dplyr` and `ggplot`) as much as possible when answering the questions. 

### Import the Dataset

Use `read_csv` to import the data, then `glimpse` and `head` to display the dataset structure.

```{r loaddatasets}
# Put your code here.
```

### Exploratory Analysis 1

#### What genders are present in the data?

Create a frequency table and barplot for the variable `gender`. 

```{r}
# Put your code here.
```

For exploration, we can do a basic barplot. The heights of the bars should match the table. Notice there are very few people (so far) who have self-identified anything besides `male` or `female`.

```{r}
# Put your code here.
```

For presentation, we should give the plot an informative title and axis labels. Adding color is optional, but it can be nice to have. Try to keep colorblind-safe options in mind if you are using colors to convey information.

```{r}
# Put your code here.
```

#### What semesters are present in the data?

Create a frequency table and barplot for the variable `semester`. 

The code is the same as what we used above, just change the variable. Notice that the semesters are in alphabetical order, which is `R`'s default for character vectors.

```{r}
# Put your code here.
```

#### Putting the semesters in correct order

Make `semester` a `factor` with the categories in the correct order. Then make frequency table and barplot using this new variable.

Academic years go spring, summer, fall. Put the years in ascending order.

```{r}
# Put your code here.
```

### Exploratory Analysis 2

#### Plots and summaries for `height`

Create a histogram, density plot, and boxplot for the variable `height`. Also calculate skewness, kurtosis, and the upper and lower fences.

This shows several `bin` options for the histogram. You often have to try several options to get a good representation. These data seem to have more than one mode (peak or local maxima)---either two or three, judging by the histogram and density plots.

```{r}
# Put your code here.
```

We can also add color and labels to histograms.

```{r}
# Put your code here.
```

The other plots follow the same pattern. The variable the same (same `aes`), just change the type of plot (different `geom`).

```{r}
# Put your code here.
```

We can calculate all four statistics in one `summarize` statement. Refer to the entries for skewness and kurtosis in the glossary for more information. There is missing data, so we need to use `na.rm = TRUE` to ignore these `NA` values.

```{r}
# Put your code here.
```

#### Plots and summaries for `armspan`

Create a histogram, density plot, and boxplot for the variable `armspan`. Also calculate skewness, kurtosis, and the upper and lower fences for the boxplot.

```{r}
# Put your code here.
```

#### Plots and summaries for `ideal`

Create a histogram, density plot, and boxplot for the variable `ideal` height. Also calculate skewness, kurtosis, and the upper and lower fences for the boxplot.

```{r}
# Put your code here.
```

### Exploratory Analysis 3

#### Filter the datasets for `gender`

Filter the data so it contains only males and females, since all other categories have too few observations for meaningful analysis. Hint: Use a filter with `%in%` to get the groups you want.

```{r}
# Put your code here.
```

#### Faceted histogram

Create a histogram for `height` faceted by `gender`, with the two plots arranged one above the other (so they will share a common x-axis scaling  the final plot). Do the same for `armspan`.

Since histograms already have two quantitative axes (the variable on `x` and the count on `y`), we need to `facet` if we want to compare groups. Faceting is similar to a `group_by()` for plots. The `facet_grid()` lets us specify a matrix or "grid" arrangement for plots using a row variable and a column variable, as `facet_grid(rowvar ~ colvar)`. Using `gender` as the row variable means the plots will have the same x-axis and same bins. We only want to facet by one variable, so we replace the other with a `.`

Here I added some axis scaling so we can see the measurements more clearly. The `breaks` are the locations of the axis tick marks. Notice what I call the "six foot bump" at about 72" in the male graph. Where do you think that comes from?

```{r}
# Put your code here.
```

#### Faceted density plot

Create a density plot for `ideal` faceted by `gender`, with the two plots arranged one above the other (so they will share a common x-axis scaling in the final plot).

Notice that the peaks are in different places on the x-axis for female and male distributions. This is why the combined distribution you explored above is bimodal.

```{r}
# Put your code here.
```

#### Comparative boxplot

Create comparative boxplots for `height` split by `gender`. Also calculate the upper and lower fences for each gender. Do the same for `ideal` height.

Boxplots only have one quantitative axis, so we can group by specifying a categorical variable for the other axis. Boxplots may be horizontal or vertical. Making them horizontal keeps the x-axis consistent with the other plots.

```{r}
# Put your code here.
```

### Concluding Discussions

#### Overall distribution shapes 

What are the overall shapes for the distributions of `height`, `armspan`, and `ideal`? Are they all unimodal? Are there any outliers?

#### Comparing male and female

Are there differences in distribution between males and females for `height`, `armspan`, and `ideal` height in this sample? Describe those differences.

#### Representativeness

What makes this sample *not* representative of people as a whole?


```{r}
sessionInfo()
```

