---
title: "Case Study: Anthropometric Data"
author: "PUT YOUR NAME HERE"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: TRUE
    toc_depth: 4
---

<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

### Load the Package(s)

```{r loadpackages}
# Put your code here.
```

### Problem Description

In our [Modeling & Regression: Theory and Line Fitting](https://stat-jet-asu.github.io/StatisticalDataAnalysis1/Slides/ModelingRegression01.html#1) slides, we posed the question of whether people are squares, as asserted by the ancient Roman architect Vitruvius. In other words, are people's heights and armspans equal? Refer to the [description page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/anthropometric.html) for Dr. Thomley's [anthropometric dataset](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/anthropometric.csv) for context regarding the data we will use to assess this claim. These data were collected over several semesters from students in Dr. Thomley’s STT1810 and STT3850 courses. Students measured one another using tape measures and/or yardsticks. Gender was self-identified.

Before doing any modeling, it is good practice to conduct exploratory data analysis on the variables in order to assess shape and determine whether there are any outliers or other anomalies, which might affect our models.

Use `tidyverse` methods (e.g., `dplyr` and `ggplot`) as much as possible when answering the questions. 

### Import the Dataset

Use `read_csv` to import the data, then `glimpse` and `head` to display the dataset structure.

```{r loaddatasets}
# Put your code here.
```

### Exploratory Analysis 1

#### What genders are present in the data?

Create a frequency table and barplot for the variable `gender`. 

```{r}
# Put your code here.
```

#### What semesters are present in the data?

Create a frequency table and barplot for the variable `semester`. 

```{r}
# Put your code here.
```

#### Putting the semesters in correct order

Make `semester` a `factor` with the categories in the correct order. Then make frequency table and barplot using this new variable.

```{r}
# Put your code here.
```

### Exploratory Analysis 2

#### Plots and summaries for `height`

Create a histogram, density plot, and boxplot for the variable `height`. Also calculate skewness, kurtosis, and the upper and lower fences.

```{r}
# Put your code here.
```

#### Plots and summaries for `armspan`

Create a histogram, density plot, and boxplot for the variable `armspan`. Also calculate skewness, kurtosis, and the upper and lower fences for the boxplot.

```{r}
# Put your code here.
```

#### Plots and summaries for `ideal`

Create a histogram, density plot, and boxplot for the variable `ideal` height. Also calculate skewness, kurtosis, and the upper and lower fences for the boxplot.

```{r}
# Put your code here.
```

### Exploratory Analysis 3

#### Filter the datasets for `gender`

Filter the data so it contains only males and females, since all other categories have too few observations for meaningful analysis. Hint: Use a filter with `%in%` to get the groups you want.

```{r}
# Put your code here.
```

#### Faceted histogram

Create a histogram for `height` faceted by `gender`, with the two plots arranged one above the other (so they will share a common x-axis scaling  the final plot). Do the same for `armspan`.

```{r}
# Put your code here.
```

#### Faceted density plot

Create a density plot for `ideal` faceted by `gender`, with the two plots arranged one above the other (so they will share a common x-axis scaling in the final plot).

```{r}
# Put your code here.
```

#### Comparative boxplot

Create comparative boxplots for `height` split by `gender`. Also calculate the upper and lower fences for each gender. Do the same for `ideal` height.

```{r}
# Put your code here.
```

### Concluding Discussions

#### Overall distribution shapes 

What are the overall shapes for the distributions of `height`, `armspan`, and `ideal`? Are they all unimodal? Are there any outliers?

#### Comparing male and female

Are there differences in distribution between males and females for `height`, `armspan`, and `ideal` height in this sample? Describe those differences.

#### Representativeness

What makes this sample *not* representative of people as a whole?


```{r}
sessionInfo()
```

