---
title: "Case Study: Oscar Winner Ages"
author: "PUT YOUR NAME HERE"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

Load the package(s) you will need.

```{r loadpackages}
# Put your code here.
```

In January 2007, Entertainment Weekly magazine featured a [cover story](http://www.simplystreep.com/gallery/displayimage.php?album=345&pid=25019#top_display_media) about three Academy Award nominees for Best Actress (Meryl Streep, Judy Densch, and Helen Mirren), with a cover tagline that proclaimed "The Silver Foxes Rule!" This theme continued inside, with the article focusing in large part on how old these three women were, relative to previous winners, and the perceived Hollywood bias in favor of younger actresses. This discussion persists in one form or another to this day. (data: [oscarages.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/oscarages.csv))

A) Use `read_csv` to import the data, then `glimpse` and `head` to display the dataset structure.

```{r}
# Put your code here.
```

B) Use `filter` to create a dataset that contains data only for Best Actor and Best Actress winners prior to the 2007 Academy Awards ceremony. Note that the variable `ceremony` refers to the ceremony year.

```{r}
# Put your code here.
```

C) Use `summarize` (with any other necessary functions) on your new dataset to compute and display the sample sizes, means, standard deviations, five-number summaries, IQR, and the lower and upper fences for Best Actor and Best Actress winners prior to 2007.

```{r}
# Put your code here.
```

D) Create boxplots to compare the ages of Best Actor and Best Actress winners prior to 2007. Add an informative title, x and y axis labels, a caption that shows the data source, and color in the plots to improve the display.

```{r}
# Put your code here.
```

E) Use `facet_grid` create histograms to compare the age distributions of Best Actor and Best Actress winners prior to 2007. Arrange the plots so they are stacked versus side-by-side. Add an informative title, x and y axis labels, a caption that shows the data source. 

```{r}
# Put your code here.
```

F) Which actress actually won the Best Actress Oscar in 2007? How old was she? For what movie did she win? Use the `filter` and `select` functions to display the answer. You'll need to go back to the full dataset for this.

```{r}
# Put your code here.
```

G) Was the 2007 Best Actress winner an outlier compared to previous Best Actress winners? If she were a man, would she have been an outlier compared to previous Best Actor winners? Determine the answers using code.

```{r}
# Put your code here.
```

H) The 2007 article made several other statements. For example (1) For the last 10 years, the Academy hasn't awarded a Best Actress Oscar to anyone over 39; (2) For the past nine years, that age limit has hovered even lower, at 35; (3) In fact, if one of our cover subjects wins next month, she'll become the most mature recipient since 1990, when Jessica Tandy took home the gold for *Driving Miss Daisy*; (4) Since then, only one actress over 50 has won the supporting category, and that was Densch herself, for *Shakespeare in Love* in '99. Can you verify these statements using code? What are the corresponding values for the Best Actor and Best Supporting Actor categories?

```{r}
# Put your code here.
```

```{r}
sessionInfo()
```

