---
title: "Case Study: STT3850 Majors #1"
author: "PUT YOUR NAME HERE"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

Load the package(s) you will need.

```{r loadpackages}

```

Load the dataset `stt3850majors.csv`.

```{r loaddata}

```

Run `glimpse()` on the dataset you created.

```{r glimpsedata}

```

Run `head()` with `n = 10` to visualize the first 10 rows.

```{r headdata}

```

Description

WRITE A BRIEF DATASET DESCRIPTION HERE

Variables

* College: WRITE A BRIEF VARIABLE DESCRIPTION HERE
* Year:  WRITE A BRIEF VARIABLE DESCRIPTION HERE
* BannerMajor:  WRITE A BRIEF VARIABLE DESCRIPTION HERE
* CodedMajor:  WRITE A BRIEF VARIABLE DESCRIPTION HERE
* Semester:  WRITE A BRIEF VARIABLE DESCRIPTION HERE
* AY:  WRITE A BRIEF VARIABLE DESCRIPTION HERE

Make `Year` and `Coded Major` factor variables. 

```{r makefactors}
stt3850majors <- stt3850majors %>% 
  mutate(Year = factor(Year, levels = c("Freshman", 
                                        "Sophomore", 
                                        "Junior", 
                                        "Senior", 
                                        "Graduate Certificate Program")),
         CodedMajor = factor(CodedMajor, levels = c("Computer Science", 
                                                    "Mathematical Sciences", 
                                                    "Other Sciences", 
                                                    "Other"))
         )
```

Use `table()` to summarize each variable in the dataset.

```{r basetables}

```

Use `count()` to summarize each variable in the dataset.

```{r dplyrcounts}

```

Summarize counts and proportions for `CodedMajor` using `dplyr`.

```{r dplyrprops}

```

Create basic barplots for `Year` and `CodedMajor`.

```{r basicbarplots}

```


***
```{r}
sessionInfo()
```