---
title: "Case Study: STT3850 Majors #1"
author: "SOLUTIONS // Jill E. Thomley"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
```

Load the package(s) you will need.

```{r loadpackages}
library(tidyverse)
```

Load the dataset `stt3850majors.csv`.

```{r loaddata}
stt3850majors <- read_csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/stt3850majors.csv"))
```

Run `glimpse()` on the dataset you created.

```{r glimpsedata}
glimpse(stt3850majors)
```

Run `head()` with `n = 10` to visualize the first 10 rows.

```{r headdata}
head(stt3850majors, n = 10)
```

Description

Data for students in Dr. Thomley's STT3850 classes from Fall 2014 to Spring 2020.

Variables

* College: ASU college associated with student major
* Year: student class level (freshman, sophomore, etc.)
* BannerMajor: student major as it appeared in Banner
* CodedMajor: majors categorized into one of four groups
* Semester: semester (S = spring, F = Fall, Sum = summer)
* AY: academic year

Make `Year` and `Coded Major` factor variables. 

```{r makefactors}
stt3850majors <- stt3850majors %>% 
  mutate(Year = factor(Year, levels = c("Freshman", 
                                        "Sophomore", 
                                        "Junior", 
                                        "Senior", 
                                        "Graduate Certificate Program")),
         CodedMajor = factor(CodedMajor, levels = c("Computer Science", 
                                                    "Mathematical Sciences", 
                                                    "Other Sciences", 
                                                    "Other"))
         )
```

Use `table()` to summarize each variable in the dataset.

```{r basetables}
table(stt3850majors$College)
table(stt3850majors$Year)
table(stt3850majors$BannerMajor)
table(stt3850majors$CodedMajor)
table(stt3850majors$Semester)
table(stt3850majors$AY)
```

Use `count()` to summarize each variable in the dataset.

```{r dplyrcounts}
count(stt3850majors, College)
count(stt3850majors, Year)
count(stt3850majors, BannerMajor)
count(stt3850majors, CodedMajor)
count(stt3850majors, Semester)
count(stt3850majors, AY)
```

Summarize counts and proportions for `CodedMajor` using `dplyr`.

```{r dplyrprops}
stt3850majors %>% 
  count(CodedMajor) %>% 
  mutate(prop = prop.table(n))

# base code version for contrast
prop.table(table(stt3850majors$CodedMajor))
```

Create basic barplots for `Year` and `CodedMajor`.

```{r basicbarplots}
ggplot(stt3850majors, aes(x = Year)) +
  geom_bar()

ggplot(stt3850majors, aes(x = Year)) + 
  geom_bar(aes(y=..count../sum(..count..))) + 
  labs(y = "proportion")

ggplot(stt3850majors, aes(x = CodedMajor)) +
  geom_bar()

ggplot(stt3850majors, aes(x = CodedMajor)) + 
  geom_bar(aes(y=..count../sum(..count..))) + 
  labs(y = "proportion")
```

***
```{r}
sessionInfo()
```